date: 27.01.2015, last modified: 09.02.2015
author: Bernd Panassiti

Figure 3- Influence of environment on Bois noir and Hyalesthes obsoletus
========================================================


Use bquote(), as in                                                                   
  n <- 200                                                                            
  vent <- bquote(paste("R con ventanas de ", .(n), " ", italic(datos)))               
                                                                                      
or, using ~ instead of paste(),                                                       
  vent <- bquote("R con ventanas de " ~ .(n) ~ italic(datos))                         
  plot(1, 1, main=vent)                                                               
                                                                                      
The notation .(xxx) means to replace xxx by the value of the variable called "xxx".   



```{r setup}
library(knitr) # for creating html-file
opts_knit$set(root.dir='../') # definining working directory; or normalizePath('../')
```

```{r load}
rm(list=ls(all=TRUE))

# load settings, data & libraries
source("r-code/R - 00 - settings.r")
load("data/predictors_response2012_work.RData")

library(rstan) # extraction of posterior, function:stanGraphInput
set_cppo("fast")  # for best running speed
library(MCMCpack) # rwish function, to create mmcmc list, function:stanGraphInput
library(modeest)
library(IDPmisc)
library(plyr) # for coda function
library(ggmcmc) # extraction of samples

set.seed(2013)
```

#### BN
```{r BN_graph_input}
load(file="results/4_rstan/output_rstan_bn_frequentgrapes.RData")
model.name="rstan_bn_frequentgrapes" # for graphic

# load posterior and disease observations
load("data/predictors_response2012_final_input_newdata.RData") # newdata
Grapes = newdata[3:(3-1+length(grapeselection))]
Environment = newdata[(3+length(grapeselection)):(3-1+length(grapeselection)+length(enviroselection))]

# stanGRAPH defaults: 
# q=0.1 (0.1 -> 0.1 & 0.)
# parGrep="par" depending how parameters are specified in Stan model, e.g.: parEnvi[1],parEnvi[2], parGrap[1],...etc

GraphInput<-stanGraphInput(stan.fit)
GraphInputBn<-GraphInput# backup
```


```{r graph_BN}
### X axis labels
abbr<-read.table("data/abbreviations.csv",sep="\t") # load correct notation
ParameterNames<-colnames(newdata)[-c(1,2,(ncol(newdata)-3):ncol(newdata))]   # extract variable names from modelling input dataframe
nameReplace<-which(ParameterNames %in% abbr[,1])    # which need to be replaced with correct notation?
for (i in nameReplace){ ParameterNames[i]<-as.vector(abbr[which(abbr[,1] %in% ParameterNames[i]),2])}
ParameterNames<-c("intercept",ParameterNames)


########### ADJUST BY HAND ###############
# which parameters should be plotted together in one graph?
parplot<-2:(length(Grapes)+1) # Grapes
parType<-"Grapes"
ylim=c(-200,20)
at=c(-200,-150,-100,-50,0,10)
mar=c(8,5,1,1)
parusrMINUSx=2

parplot<-c(1,(length(Grapes)+2):(length(Grapes)+1+length(Environment))) # Environment
parType<-"Environment"
xlim=c(-10,5)
at=c(-15,-10,-5,0,5,10)
mar=c(17,6,1,1)
parusrMINUSx=0.5
n.ch=4
###########################################


# MEDIANS OF RUNS
var.median.plot<-GraphInput$var.median[GraphInput$var.median[,1] %in% parplot,]
var.median.plot[,1]<-rep(1:length(parplot),each=n.ch)

# CREDIBLE INTERVAL
datq.plot<-GraphInput$datq[GraphInput$datq[,1] %in% parplot,]
datq.plot[,1:2]<-1:length(parplot)

# RANKING
# add mean of the four medians for ranking
datq.plot$mean <- 0
for (i in 1:nrow(datq.plot)){
  datq.plot[i,5] <- mean(var.median.plot[var.median.plot[,1]==i,2])
}

# delete intercept
datq.plot <- datq.plot[-1,]

var.median.plot$parameter <- rep(ParameterNames[parplot],each=4)
var.median.plot <- var.median.plot[-c(1:4),]

# rank & order according ranking
datq.plot$mean_rank<-rank(datq.plot[,5])                    # ranking credible interval
var.median.plot$mean_rank<-rep(rank(datq.plot[,5]),each=4)  # ranking median

df                <- datq.plot[order(datq.plot[,6]),]             # sort for mean_rank
var.median.qqplot <- var.median.plot[order(var.median.plot[,4]),] # sort for mean_rank


# preparation of data.frame for qqplot
datq.qqplot<-data.frame(matrix(0,nrow=nrow(df)*2,ncol=4))
j=1
for (i in 1:nrow(df)){
  datq.qqplot[j,c(1:2,3,4)]  <- df[i,c(2,3,5,6)]
  datq.qqplot[j+1,c(1:2,3,4)]  <- df[i,c(2,4,5,6)]
j=j+2
}

colnames(datq.qqplot)       <-c("id","credible_range","mean","mean_rank")
colnames(var.median.qqplot) <-c("id","median","parameter","mean_rank")

# save
BNdatq.qqplot <- datq.qqplot
BNvar.median.qqplot <- var.median.qqplot
datq.qqplot <- BNdatq.qqplot
var.median.qqplot <- BNvar.median.qqplot

# y labels
yLabels <- unique(var.median.qqplot$parameter)
yLabels2 <- yLabels
yLabels2[11] <- expression(paste(italic(Hyalesthes)," ",italic(obsoletus)," ","predictions"))
qp <- ggplot(data=datq.qqplot, aes(x=credible_range, y=mean_rank, group=mean_rank)) + geom_line()
BNqqplot <- qp + 
  geom_point(data=var.median.qqplot,aes(x=median,y=mean_rank)) +
  scale_y_discrete(breaks=unique(var.median.qqplot$mean_rank),labels=yLabels2) +
  geom_vline(xintercept=0,size=0.2)+
  theme_bw() + # removes grey background
  theme(axis.line = element_line(colour='black',size=0.5),
        axis.title = element_blank() # remove axis labels
        )
```


### Ho
```{r Ho_graph_input}
rm(stan.fit) # delete BN stan.fit
rm(newdata)  # delete newdata - input for BN model
load(file="results/4_rstan/output_rstan_ho_100000.RData") # contains stan.fit and newdata


enviroselection = c(4:14)
Environment = newdata[3:(3-1+length(enviroselection))]


GraphInput<-stanGraphInput(stan.fit)
GraphInputHo<-GraphInput # backup
```

```{r Ho_graph_preparation}
######### ADJUST BY HAND ###########
model.name="rstan_ho" # for graphic
n.ch=4 # number of chains
n.iter.total=100000
n.draws<-100000/100 # total samples divided by final samples size
n.draws.new=1000
nObs=length(newdata$ill)

### X axis labels
abbr<-read.table("data/abbreviations.csv",sep="\t") # load correct notation
ParameterNames<-colnames(newdata)[-c(1,2)]   # extract variable names from modelling input dataframe
nameReplace<-which(ParameterNames %in% abbr[,1])    # which need to be replaced with correct notation?
for (i in nameReplace){ ParameterNames[i]<-as.vector(abbr[which(abbr[,1] %in% ParameterNames[i]),2])}
ParameterNames<-c("intercept",ParameterNames)



########### ADJUST BY HAND ###############
# which parameters should be plotted together in one graph?
parplot<-1:(length(Environment)+1) # Environment
parType<-"Environment"
yupper=230
ylower=-120
ylim=c(ylower,yupper)
at=seq(ylower,yupper,50) # numeric vector where ticks should be drawn
parusrMINUSx=20
mar=c(17,6,1,1)
###########################################


var.median.plot<-GraphInput$var.median[GraphInput$var.median[,1] %in% parplot,]
var.median.plot[,1]<-rep(1:length(parplot),each=n.ch)
datq.plot<-GraphInput$datq[GraphInput$datq[,1] %in% parplot,]
datq.plot[,1:2]<-1:length(parplot)


# RANKING
# based on "BNvar.median.qqplot"
datq.plot$mean <- 0
for (i in 1:nrow(datq.plot)){
  datq.plot[i,5] <- mean(var.median.plot[var.median.plot[,1]==i,2])
}

# delete intercept
#datq.plot <- datq.plot[-1,]

datq.plot$parameter <- ParameterNames[parplot] # assign variable names to intervals
var.median.plot$parameter <- rep(ParameterNames[parplot],each=4)
#var.median.plot <- var.median.plot[-c(1:4),] # when intercept is not needed

# rank & order according ranking
# variables ranking are synchronized with that from BN model

# datq.plot$mean_rank<-rank(datq.plot[,5])                    # ranking credible interval

# > datq.plot
#    X1 X2          X3          X4       mean                          parameter
# 1   1  1 -3.71570560 23.24714997 10.0893622                          intercept
# 2   2  2 -8.84270941  0.88758754 -3.5809379           cation exchange capacity
# 3   3  3 -6.10586649  0.84009127 -2.6122812          precipitation seasonality
# 4   4  4  1.33634875  9.52672049  4.1120640                      isothermality
# 5   5  5 -3.15947761 10.96274256  2.5390761  min temperature of coldest period
# 6   6  6 -6.70536454  1.76097134 -1.1979371 mean temperature of driest quarter
# 7   7  7 -7.39594407  2.85838340 -1.2646789            mean annual temperature
# 8   8  8 -5.92716716  8.68756929  2.2921944                            soil pH
# 9   9  9  8.91648709 23.62616175 16.1291624                 natural vegetation
# 10 10 10 -9.63668545  7.25690355 -0.7802193           effective field capacity
# 11 11 11  3.45714061 15.33539478  8.8162528                       soil erosion
# 12 12 12 -0.09866019 -0.02768301 -0.0627370                    altitude a.s.l.


# # exposition is missing in Ho model -> ADDITIONAL LINE(s) # #
datq.plot <- rbind(datq.plot,rep(0,ncol(datq.plot)))
datq.plot[nrow(datq.plot),6] <- ""

var.median.plot <- rbind(var.median.plot,rep(0,ncol(var.median.plot)),rep(0,ncol(var.median.plot)),rep(0,ncol(var.median.plot)),rep(0,ncol(var.median.plot)))

var.median.plot[(nrow(var.median.plot)-3):nrow(var.median.plot),3] <- ""
#############################################################################
# # intercept not available in BN model -> gets rank of Ho predictor
datq.plot$mean_rank<-c(11,2,7,5,1,4,12,9,8,10,6,13,3)

var.median.plot$mean_rank<-rep(datq.plot[,7],each=4)  # ranking median

df                <- datq.plot[order(datq.plot[,7]),-6]           # sort for mean_rank
var.median.qqplot <- var.median.plot[order(var.median.plot[,4]),] # sort for mean_rank


# preparation of data.frame for qqplot
datq.qqplot<-data.frame(matrix(0,nrow=nrow(df)*2,ncol=4))
j=1
for (i in 1:nrow(df)){
  datq.qqplot[j,c(1:2,3,4)]  <- df[i,c(2,3,5,6)]
  datq.qqplot[j+1,c(1:2,3,4)]  <- df[i,c(2,4,5,6)]
  j=j+2
}

colnames(datq.qqplot)       <-c("id","credible_range","mean","mean_rank")
colnames(var.median.qqplot) <-c("id","median","parameter","mean_rank")

# save
HOdatq.qqplot <- datq.qqplot
HOvar.median.qqplot <- var.median.qqplot

# y labels
yLabels <- unique(var.median.qqplot$parameter)

```

```{r Ho_graph}
qp <- ggplot(data=datq.qqplot, aes(x=credible_range, y=mean_rank, group=mean_rank)) + geom_line()
HOqqplot <- qp + 
  geom_point(data=var.median.qqplot,aes(x=median,y=mean_rank)) +
  scale_y_discrete(breaks=unique(var.median.qqplot$mean_rank),labels=yLabels) +
  geom_vline(xintercept=0,size=0.2)+
  theme_bw() + # removes grey background
  theme(axis.line = element_line(colour='black',size=0.5),
        axis.title = element_blank() # remove axis labels
        )
```


### Combine Bn and Ho conditional distributions in one graph
```{r}
Layout <- grid.layout(nrow = 1, ncol = 2, 
                      widths = unit(c(2,2), c("null", "null")), heights = unit(c(1,1), c("null", "null")))
vplayout <- function(...) {
  grid.newpage()
  pushViewport(viewport(layout = Layout))
}
subplot <- function(x, y) viewport(layout.pos.row = x,
                                   layout.pos.col = y)



vplayout()
print(BNqqplot, vp = subplot(1, 1))
print(HOqqplot, vp = subplot(1, 2))
```


### 
```{r save}
save(
  GraphInputBn,GraphInputHo,
  BNdatq.qqplot,BNvar.median.qqplot,BNqqplot,
  HOdatq.qqplot,HOvar.median.qqplot,HOqqplot,
  file="results/4_rstan/output_rstan_graphs.RData")
```

