---
title: "Environmental drivers of white plague disease on shallow and mesophotic coral reefs, U.S. Virgin Islands"
author: 
-name:"Andia Chaves Fonnegra and Bernd Panassiti"
  affiliation: University of the Virgin Islands (US.VI) and Laimburg Research Center (Italy)
  abstract: "This document takes the the raw data file (.csv) and save it as as .rdata used on the article:"Environmental drivers of white plague disease on shallow and mesophotic coral reefs, U.S. Virgin Islands"."
  keywords: rawdata
date: '08.23.2016; last modified: 07.25.2017 `r format(Sys.Date(), format="%d.%m.%Y")`'
output: 
  pdf_document: 
    number_sections: yes
    toc: yes
header-includes: \usepackage{graphicx}
email: \email{}
---

```{r setup, include=FALSE, warnings=FALSE}
library(knitr)
opts_knit$set(root.dir='../')                     # definining working directory; or normalizePath('../')
opts_chunk$set(fig.align='center',                # aligns all figures
                echo=TRUE,                        # shows r-code
                message=FALSE,                    # suppresses library outputs
                warnings=FALSE,                   # suppresses library outputs
                tidy=TRUE,  # prevents the source code from running off a pdf page
                dev='pdf')                        # pdf device
```

# Introduction
Compilation of raw data from the csv file compiled and prepared by Andia Chaves Fonnegra based on several excel files that included disease data and environmental and biological data. Disease data were obtained by Marilyn Brandt, and Biological data obtained by Tyler Smith. Oceanographic-Environmental Data collected in the field by Marilyn Brandt and Tyler Smith, and downloaded and cleaned in the lab by Vanessa Wrigh and Jonathan Jossart. Loading Raw data file without average of ill a ntot of White Plague - made July 11-. Rainfall was changed to mm and rows with average data for diseased (ill or ntot) was eliminated , this to avoid adding more variability as we discussed with Bernd)

# Data

##Loading data 
```{r}
rm(list=ls(all=TRUE))
# Load settings

# load data & libraries 
Whiteplague_rawdata <- read.csv("data/Rawdata_WhitePlague.csv", header = TRUE) 

#check data
ls(Whiteplague_rawdata)
dim(Whiteplague_rawdata)
head(Whiteplague_rawdata)

#Save data as rdata file to avoid changing this data in the future
save(Whiteplague_rawdata,file="data/Whiteplague_rawdata.Rdata")
```

Once this file is saved, we used in the Whiteplague_working data Rmarkdown script to modify it and scaled/ transformed data without
changing the original rawdata.


