---
title: "linear model disease patters"
author: "Andia Chaves Fonnegra"
date: "5/28/2018"
output: pdf_document
---

```{r setup, include=FALSE, warnings=FALSE}
library(knitr)
opts_knit$set(root.dir='../')                     # definining working directory; or normalizePath('../')
opts_chunk$set(fig.align='center',                # aligns all figures
                echo=TRUE,                        # shows r-code
                message=FALSE,                    # suppresses library outputs
                warnings=FALSE,                   # suppresses library outputs
                tidy=TRUE,  # prevents the source code from running off a pdf page
                dev='pdf')                        # pdf device
```

#Introduction

This document provides the r-code for a linaer model to understand the patterns of white plague disease on St. Thomas coral reefs. The specific objectives of this analysis are: (1) to evaluate if patterns of disease are different among depth, and (2) to evaluate how disease patterns are per month, season and year). 

#Data
The white plague prevalence data is teh same we sued for the rstan biological and environmental models, cna be dowloaded inxxxxxxx. The file "whiteplague_working.RData" needs to be placed in a folder called "data".
"


## Load data
```{r}
rm(list=ls(all=TRUE))


load(file="data/Whiteplague_workingdata.Rdata")
head(Whiteplague_workingdata)
dim(Whiteplague_workingdata)

WP<-Whiteplague_workingdata
head(WP)
```


#Check probability distribution in the data.
# I follow suggestions by: http://ase.tufts.edu/gsc/gradresources/guidetomixedmodelsinr/mixed%20model%20guide.html
#The y axis represents the observations and the x axis represents the quantiles modeled by the distribution. 
#The solid red line represents a perfect distribution fit and the dashed red lines are the confidence intervals of the perfect distribution fit. 

```{r}
#normality test
#Shapiro-Wilkinson normality test
shapiro.test(WP$Plague)
#for producing a normal quantile-quantile plot
qqnorm(WP$Plague)
qqline(WP$Plague)

#If p> than 0.05. Data is normal, if not, not normal
#WP data is not normal. However it folowws a log normal
#Bartlest test of homegenty of variances
bartlett.test(WP$Plague~WP$Station_code)
#If p> than 0.05. variances are homogeneous, if not, not homogeneous

```

#check outliers and data in general with boxplot
```{r}
par(mfrow=c(2,2))
plot(WP$Plague)
boxplot(WP$Plague)
hist(WP$Plague,main="")
y2<-WP$Plague
y2[52]<-21.75
plot(y2)

summary(WP$Plague)
```


#Checking which distribution white plague data follows.
#You want to pick the distribution for which the largest number of #observations falls between the dashed lines. 
```{r}
## Loading required package: car
library(car)
library(MASS)
# This is to check if follows a normal distribution. But we need to ad the value of 1 to avoid problems
Plague<-WP$Plague+1

qqp(Plague, "norm")
# This is to check if follows a LOG normal distribution
qqp(Plague, "lnorm")

#In this case, that's the #lognormal distribution, is the best distribution for the data we have.
```


#Now, armed with the knowledge of which probability distribution fits best, I can try fitting a model.

#I choose a linear mix effect model (lmer). lmer fits as  a linear #mixed-effects models with specified mixtures of fixed effects and
#random effects and allow for the specification of correlation structure #amongst the explanatory variables and autocorrelation of the response #variable (e.g. time series effects with repeated measures). lmer allows for #non-normal errors and non-constant variance with the same error families as a GLM.

#So data is not normal and has autocorrelation due to the time series effect with repeated measurement.
#------------------------------------
#Some information from the R book:
#Deciding the fix and random effects to undertand patterns of white plagued disease. Mixed-effects models are so called because the explanatory variables are a mixture of fixed effects and random effects:
#• fixed effects influence only the mean of y;
#• random effects influence only the variance of y.

#Fixed effects are unknown constants to be estimated from the data. Random effects govern the variance – covariance structure of the response variable. #

#The fixed effects are often experimental treatments that
#were applied under our direction.

#The random effects are either categorical or continuous variables that are #distinguished by the fact that we are typically not interested in the #parameter values, but only in the variance they explain.
#-----------------------------------------------

#Pick fixed and random effects
```{r}
#Fixed effects are station and depth (experimental treatment)
#Random effects are month, season and year (temporal pseudoreplication)

#Random effect needs to indicate that the monthly measurement 
#represents pseudoreplication within each station
```


#For data inspections we start with a pannel plot also called trellis plots or lattice plots for that we need nlme and lattice libraries
```{r}
library(lattice)
require(lme4)
library(lme4)
```

#First we group the dataframe into a groupedData object. Here we indicate the fixed effect by defining temperature as outer 
#Here we get plot for each Site (Station)

```{r}
#here I indicate that depth and site are fixed effects that interact (B), whereas month, season and year are random (A) Simultaneously the intercept and the slope of B depend of time (A)
#lmer(Y ~ 1 + B + (1 + B | A), data=d)

#test wp per year
model1 <- lmer(WP$Plague~1+(1|WP$Year),REML = FALSE)
summary(model1)
qqp(residuals(model1),"lnorm")

#adding depth as fixed factor, and depth as depending of month
model2 <- lmer(WP$Plague~WP$Depth_transect + (1+WP$Depth_transect|WP$Month_12),REML = FALSE)
summary(model2)
qqp(residuals(model2),"lnorm")
#good fit

#using station o as fixed factor and months
model3 <- lmer(WP$Plague~WP$Station_code + (1+WP$Station_code|WP$Month_12), REML = FALSE)
summary(model3)
qqp(residuals(model3),"lnorm")
#did not converged

#adding depth as fixed factor, and depth as depending of year
model4 <- lmer(WP$Plague~WP$Depth_transect + (1+WP$Depth_transect|WP$Year),REML = FALSE)
               
summary(model4)
qqp(residuals(model4),"lnorm")


#adding depth as fixed factor, and depth as depending of year
model5 <- lmer(WP$Plague~WP$Depth_transect + (1+WP$Depth_transect|WP$Season), REML = FALSE)
               
summary(model5)
qqp(residuals(model5),"lnorm")


#real model taking into account all fixed and random factors is not possible. It does not allowed to nest month, seasons and years
modelall <- lmer(WP$Plague~WP$Depth_transect + (1+WP$Depth_transect|WP$Year/WP$Season/WP$Month_12))
               summary(modelall)
qqp(residuals(modelall),"lnorm")
#So it is asking for a variable that group time factor as one, basically a variable that combines Year-Season and Month.

```

#Create a new time variable combining month, season and year
```{r}
WP$time <-as.character(interaction(WP$Year,WP$Season,WP$Month_12))

WP$time
head(WP)

```

#Run model with the time variable as a random variable
```{r}
#real model taking into account depth as fixed factor and time as random (time variable is a combined variable that including month, seasons and years)
modelall <- lmer(WP$Plague~WP$Depth_transect + (1+WP$Depth_transect|WP$time),REML = FALSE )
summary(modelall)
qqp(residuals(modelall),"lnorm")
#So it is asking for a variable that group time factor as one, basically a variable that combines Year-Season and Month.
```

#Binomial models
#Generalized Mixed-effect model using Depth, Month, Season and Year as fixed factors and accounting for the temporal and spatial pseudoreplication using a random factor variable (magic variable that is a combination of Year-Season-Month-Site)

#Libraries
```{r}
library(lme4)
library(glmm)
library(coefplot) ## for coefplot2
library(reshape)
library(plyr)
library(ggplot2)
library(gridExtra)
library(emdbook) ## for qchibarsq
library(MASS)

```


#Binomial model
```{r}
attach(WP)

model1<- glmer(Plague ~ Depth_transect + (1 | time), family="binomial"(link="logit"),data=WP, nAGo=25)

summary(model1)
qqp(residuals(model1),"lnorm")

```

#GLMM to test WP disease pattersn in relation to depth, month, season and year.
```{r}
# Using GLM (family= Binomial) # for White Plague Pervalence in relation to other environmental variables
str(WP)
#Check all variables as factors
#Evaluating WP prevalence in relation to Depth and time
WP$Depth_transect <-as.factor(WP$Depth_transect) # Change Depth as factor
WP$Year<-as.factor(WP$Year) # Change Year as factor
WP$Season<-as.factor(WP$Season) # Change Season as factor
WP$Month_12<-as.factor(WP$Month_12) # Change Month as factor
WP$time<-as.factor(WP$time)


#Give number of corals with and without diseases for the binomial, as it needs both to run
not.ill<- WP$Ntot-WP$ill
not.ill
ill<-WP$ill
ill
y<-cbind(ill,not.ill) # binding both vectors of ill and no.ill corals - binomial

#This is a simplify example with glm)
m1<- glm(y~WP$Depth_transect,family= binomial)
print(summary(m1))
qqp(residuals(m1),"norm")
#---------------------------------------------

#Now with glmm------------

#just Depth and using Month to account for the temporal pseudoreplication
m2<- glmer(y ~ Depth_transect + (1 | Month_12), family="binomial"(link="logit"),data=WP)
print(summary(m2))
qqp(residuals(m2),"norm")

#Just Depth and using time to account for the temporal pseudoreplication
m3<- glmer(y ~ Depth_transect + (1 | time), family="binomial"(link="logit"),data=WP)
print(summary(m3))
qqp(residuals(m3),"norm")
#--results as same with only month---

#Just Depth and using depth linked time to account for the spatial and temporal pseudoreplication
m4<- glmer(y ~ Depth_transect + (1 + Depth_transect| time), family="binomial"(link="logit"),data=WP)
print(summary(m4))
qqp(residuals(m4),"norm")
#--Not possibleE
#Error: number of observations (=239) < number of random effects (=260) for term (1 + Depth_transect | time); the random-effects parameters are probably unidentifiable#

#So, adding in the random effecr month as nested within depth for account for pseudoreplication.
m5<- glmer(y ~ Depth_transect + (1| Depth_transect/Month_12), family="binomial"(link="logit"),data=WP)
print(summary(m5))
qqp(residuals(m5),"norm")
#--this one improve residuals fitting

#Depth as fixed, change in random for station/time. Station works better
m6<- glmer(y ~ Depth_transect + (1|Station_code/time), family="binomial"(link="logit"),data=WP)
print(summary(m6))
qqp(residuals(m6),"norm")

##Depth as fixed, change in random for station/Month.Results evry similar to m6
m7<- glmer(y ~ Depth_transect + (1|Station_code/Month_12), family="binomial"(link="logit"),data=WP)
print(summary(m7))
qqp(residuals(m7),"norm")

#Combining both depth and month as fixed factors 
m8<- glmer(y ~ Depth_transect*Month_12 + (1|Station_code/Month_12), family="binomial"(link="logit"),data=WP)
print(summary(m8))
qqp(residuals(m8),"norm")
#there is not effect of month when adding the depth interaction

#Only checking for month effect
m9<- glmer(y ~ Month_12 + (1|Station_code/Month_12), family="binomial"(link="logit"),data=WP)
print(summary(m9))
qqp(residuals(m9),"norm")
#there is an effect per month.

#Only season
m9.1<- glmer(y ~ Season + (1|Station_code/Month_12), family="binomial"(link="logit"),data=WP)
print(summary(m9.1))
qqp(residuals(m9.1),"norm")


#Season and depth 

m10<- glmer(y ~ Depth_transect*Season +(1|Station_code/Month_12), family="binomial"(link="logit"),data=WP)
print(summary(m10))
qqp(residuals(m10),"norm")
#there is 

#Only Season
m11<- glmer(y ~ Season +(1|Station_code/Month_12), family="binomial"(link="logit"),data=WP)
print(summary(m11))
qqp(residuals(m11),"norm")


#Only Year
m11<- glmer(y ~ Depth_transect*Year +(1|Station_code/time), family="binomial"(link="logit"),data=WP)
print(summary(m11))
qqp(residuals(m11),"norm")

#All, glmer(Y ~ D*M + D*S + D*Y + (1|Station/time(Y-M-S)), data=d
m12<- glmer(y ~ Depth_transect*Month_12+Depth_transect*Season+Depth_transect*Year +(1|Station_code/time), family="binomial"(link="logit"),data=WP)
print(summary(m12))
qqp(residuals(m12),"norm")


#All, glmer(Y ~ D + M + S +Y + (1|Station/time(Y-M-S)), data=d
m12<- glmer(y ~ Depth_transect+Season+Month+Year +(1|Station_code/time), family="binomial"(link="logit"),data=WP)
print(summary(m12))
qqp(residuals(m12),"norm")


```

