---
title: "Graphs, GLMS, Correlations-Checking Data"
author: "Andia Chaves Fonnegra"
date: "7/17/2017"
output: pdf_document
---

```{r setup, include=FALSE, warnings=FALSE}
library(knitr)
opts_knit$set(root.dir='../')                     # definining working directory; or normalizePath('../')
opts_chunk$set(fig.align='center',                # aligns all figures
                echo=TRUE,                        # shows r-code
                message=FALSE,                    # suppresses library outputs
                warnings=FALSE,                   # suppresses library outputs
                tidy=TRUE,  # prevents the source code from running off a pdf page
                dev='pdf')                        # pdf device
```

# Data

# Joint model
## Loading data 
```{r}
rm(list=ls(all=TRUE))
# Load settings
source('r-code/00_settings.R')


# load data & libraries - Loading Raw data file without average of ill a ntot of White Plague - made July 11-Andia 
load(file="data/Whiteplague_workingdata.Rdata")

library(ggplot2) 
```


Boxplots - figures
```{r}
White <- Whiteplague_workingdata


#BOX-PLOTS PER MONTH ENVIRONMENTAL VARIABLES
#Chlorophyll
White$Month_12<-factor(White$Month_12)  #Convert Month into a labelled factor to use as a grouping variabel
p1 <- ggplot(White, aes(x = Month_12, y = Chlorophyll))+ scale_y_continuous(name = "Chlorophyll"~("mg m"^{-3})) + scale_x_discrete(name = "Months") +
      geom_boxplot()
p1


#Conductivity
White$Month_12<-factor(White$Month_12)  #Convert Month into a labelled factor to use as a grouping variabel
p2 <- ggplot(White, aes(x = Month_12, y = Conductivity)) +scale_y_continuous(name = "Conductivity"~("COS m"^{-1})) + scale_x_discrete(name = "Months") +
      geom_boxplot()
p2

#Density
White$Month_12<-factor(White$Month_12)  #Convert Month into a labelled factor to use as a grouping variabel
p3 <- ggplot(White, aes(x = Month_12, y = Density)) +scale_y_continuous(name = "Density"~("Kg"%*%"m"^{-3})) + scale_x_discrete(name = "Months") +
        geom_boxplot()
p3

#MAX DHW
White$Month_12<-factor(White$Month_12)  #Convert Month into a labelled factor to use as a grouping variabel
p4 <- ggplot(White, aes(x = Month_12, y = Max_DHW)) +  scale_y_continuous(name = "Max DHW") + scale_x_discrete(name = "Months") + geom_boxplot()
p4

#Oxygen
White$Month_12<-factor(White$Month_12)  #Convert Month into a labelled factor to use as a grouping variabel
p5 <- ggplot(White, aes(x = Month_12, y = Oxygen)) + scale_y_continuous(name = "Dissolved Oxygen"~("mg L"^{-1})) + scale_x_discrete(name = "Months") + 
     geom_boxplot()
p5

#Rainfall
White$Month_12<-factor(White$Month_12)  #Convert Month into a labelled factor to use as a grouping variabel
p6 <- ggplot(White, aes(x = Month_12, y = Rainfall)) + scale_y_continuous(name = "Rainfall (inches)") + scale_x_discrete(name = "Months") + 
geom_boxplot()
p6

#Salinity
White$Month_12<-factor(White$Month_12)  #Convert Month into a labelled factor to use as a grouping variabel
p7 <- ggplot(White, aes(x = Month_12, y = Salinity)) + scale_y_continuous(name = "Salinity (PSU)") + scale_x_discrete(name = "Months") + 
geom_boxplot()
p7

#Temperature
White$Month_12<-factor(White$Month_12)  #Convert Month into a labelled factor to use as a grouping variabel
p01 <- ggplot(White, aes(x = Month_12, y = Temperature)) + scale_y_continuous(name = "Temperature"~(degree*C)) + scale_x_discrete(name = "Months") + 
     geom_boxplot()
p01

#Turbidity
White$Month_12<-factor(White$Month_12)  #Convert Month into a labelled factor to use as a grouping variabel
p9 <- ggplot(White, aes(x = Month_12, y = Turbidity)) +  scale_y_continuous(name = "Turbidity (NTU)") + scale_x_discrete(name = "Months") + 
geom_boxplot()
p9

#WP prevalence
White$Month_12<-factor(White$Month_12)  #Convert Month into a labelled factor to use as a grouping variabel
p10 <- ggplot(White, aes(x = Month_12, y = White_Plague_Pervalence)) + scale_y_continuous(name = "WP Prevalence (diseased/total)") + scale_x_discrete(name = "Months") + geom_boxplot()
p10


```

BOX PLOTS ENVIRONMENTAL PER SEASON

```{r}
#BOX PLOTS ENVIRONMENTAL PER SEASON
#Chlorophyll
White$Season<-factor(White$Season)  #Convert Month into a labelled factor to use as a grouping variabel
p11 <- ggplot(White, aes(x = Season, y = Chlorophyll))+ scale_y_continuous(name = "Chlorophyll"~("mg m"^{-3})) + scale_x_discrete(name = "Seasons") +
      geom_boxplot()
p11


#Conductivity
White$Season<-factor(White$Season)  #Convert Month into a labelled factor to use as a grouping variabel
p12 <- ggplot(White, aes(x = Season, y = Conductivity)) +scale_y_continuous(name = "Conductivity"~("COS m"^{-1})) + scale_x_discrete(name = "Seasons") +
      geom_boxplot()
p12

#Density
White$Season<-factor(White$Season)  #Convert Month into a labelled factor to use as a grouping variabel
p13 <- ggplot(White, aes(x = Season, y = Density)) +scale_y_continuous(name = "Density"~("Kg"%*%"m"^{-3})) + scale_x_discrete(name = "Seasons") +
        geom_boxplot()
p13

#MAX DHW
White$Season<-factor(White$Season)  #Convert Month into a labelled factor to use as a grouping variabel
p14 <- ggplot(White, aes(x = Season, y = Max_DHW)) +  scale_y_continuous(name = "Max DHW") + scale_x_discrete(name = "Seasons") + geom_boxplot()
p14

#Oxygen
White$Season<-factor(White$Season)  #Convert Month into a labelled factor to use as a grouping variabel
p15 <- ggplot(White, aes(x = Season, y = Oxygen)) + scale_y_continuous(name = "Dissolved Oxygen"~("mg L"^{-1})) + scale_x_discrete(name = "Seasons") + 
     geom_boxplot()
p15

#Rainfall
White$Season<-factor(White$Season)  #Convert Month into a labelled factor to use as a grouping variabel
p16 <- ggplot(White, aes(x = Season, y = Rainfall)) + scale_y_continuous(name = "Rainfall (inches)") + scale_x_discrete(name = "Seasons") + 
geom_boxplot()
p16

#Salinity
White$Season<-factor(White$Season)  #Convert Month into a labelled factor to use as a grouping variabel
p17 <- ggplot(White, aes(x = Season, y = Salinity)) + scale_y_continuous(name = "Salinity (PSU)") + scale_x_discrete(name = "Seasons") + 
geom_boxplot()
p17

#Temperature
White$Season<-factor(White$Season)  #Convert Month into a labelled factor to use as a grouping variabel
p18 <- ggplot(White, aes(x = Season, y = Temperature)) + scale_y_continuous(name = "Temperature"~(degree*C)) + scale_x_discrete(name = "Seasons") + 
     geom_boxplot()
p18

#Turbidity
White$Season<-factor(White$Season)  #Convert Month into a labelled factor to use as a grouping variabel
p19 <- ggplot(White, aes(x = Season, y = Turbidity)) +  scale_y_continuous(name = "Turbidity (NTU)") + scale_x_discrete(name = "Seasons") + 
geom_boxplot()
p19

#WP prevalence
White$Season<-factor(White$Season)  #Convert Month into a labelled factor to use as a grouping variabel
p20 <- ggplot(White, aes(x = Season, y = White_Plague_Pervalence)) + scale_y_continuous(name = "WP Prevalence (diseased/total)") + scale_x_discrete(name = "Seasons") + geom_boxplot()
p20
```

BOX PLOTS ENVIRONMENTAL PER YEAR

```{r}
#Chlorophyll
White$Year<-factor(White$Year)  #Convert Month into a labelled factor to use as a grouping variabel
p21 <- ggplot(White, aes(x = Year, y = Chlorophyll))+ scale_y_continuous(name = "Chlorophyll"~("mg m"^{-3})) + scale_x_discrete(name = "Years") +
      geom_boxplot()
p21


#Conductivity
White$Year<-factor(White$Year)  #Convert Month into a labelled factor to use as a grouping variabel
p22 <- ggplot(White, aes(x = Year, y = Conductivity)) +scale_y_continuous(name = "Conductivity"~("COS m"^{-1})) + scale_x_discrete(name = "Years") +
      geom_boxplot()
p22

#Density
White$Year<-factor(White$Year)  #Convert Month into a labelled factor to use as a grouping variabel
p23 <- ggplot(White, aes(x = Year, y = Density)) +scale_y_continuous(name = "Density"~("Kg"%*%"m"^{-3})) + scale_x_discrete(name = "Years") +
        geom_boxplot()
p23

#MAX DHW
White$Year<-factor(White$Year)  #Convert Month into a labelled factor to use as a grouping variabel
p24 <- ggplot(White, aes(x = Year, y = Max_DHW)) +  scale_y_continuous(name = "Max DHW") + scale_x_discrete(name = "Years") + geom_boxplot()
p24

#Oxygen
White$Year<-factor(White$Year)  #Convert Month into a labelled factor to use as a grouping variabel
p25 <- ggplot(White, aes(x = Year, y = Oxygen)) + scale_y_continuous(name = "Dissolved Oxygen"~("mg L"^{-1})) + scale_x_discrete(name = "Years") + 
     geom_boxplot()
p25

#Rainfall
White$Year<-factor(White$Year)  #Convert Month into a labelled factor to use as a grouping variabel
p26 <- ggplot(White, aes(x = Year, y = Rainfall)) + scale_y_continuous(name = "Rainfall (inches)") + scale_x_discrete(name = "Years") + 
geom_boxplot()
p26

#Salinity
White$Year<-factor(White$Year)  #Convert Month into a labelled factor to use as a grouping variabel
p27 <- ggplot(White, aes(x = Year, y = Salinity)) + scale_y_continuous(name = "Salinity (PSU)") + scale_x_discrete(name = "Years") + geom_boxplot()
p27

#Temperature
White$Year<-factor(White$Year)  #Convert Month into a labelled factor to use as a grouping variabel
p28 <- ggplot(White, aes(x = Year, y = Temperature)) + scale_y_continuous(name = "Temperature"~(degree*C)) + scale_x_discrete(name = "Years") + 
     geom_boxplot()
p28

#Turbidity
White$Year<-factor(White$Year)  #Convert Month into a labelled factor to use as a grouping variabel
p29 <- ggplot(White, aes(x = Year, y = Turbidity)) +  scale_y_continuous(name = "Turbidity (NTU)") + scale_x_discrete(name = "Years") + 
geom_boxplot()
p29

#WP prevalence
White$Year<-factor(White$Year)  #Convert Month into a labelled factor to use as a grouping variabel
p30 <- ggplot(White, aes(x = Year, y = White_Plague_Pervalence)) + scale_y_continuous(name = "WP Prevalence (diseased/total)") + scale_x_discrete(name = "Years") + geom_boxplot()
p30
```

BOX PLOTS ENVIRONMENTAL PER DEPTH

```{r}
#Chlorophyll
White$Depth_disease_transect<-factor(White$Depth_disease_transect)  #Convert depth into a labelled factor to use as a grouping variabel
p31 <- ggplot(White, aes(x = Depth_disease_transect, y = Chlorophyll))+ scale_y_continuous(name = "Chlorophyll"~("mg m"^{-3})) + scale_x_discrete(name = "Depth (m)") +
      geom_boxplot()
p31


#Conductivity
White$Depth_disease_transect<-factor(White$Depth_disease_transect)  #Convert depth into a labelled factor to use as a grouping variabel
p32 <- ggplot(White, aes(x = Depth_disease_transect, y = Conductivity)) +scale_y_continuous(name = "Conductivity"~("COS m"^{-1})) + scale_x_discrete(name = "Depth (m)") +
      geom_boxplot()
p32

#Density
White$Depth_disease_transect<-factor(White$Depth_disease_transect)  #Convert depth into a labelled factor to use as a grouping variabel
p33 <- ggplot(White, aes(x = Depth_disease_transect, y = Density)) +scale_y_continuous(name = "Density"~("Kg"%*%"m"^{-3})) + scale_x_discrete(name = "Depth (m)") +
        geom_boxplot()
p33

#MAX DHW
White$Depth_disease_transect<-factor(White$Depth_disease_transect)  #Convert depth into a labelled factor to use as a grouping variabel
p34 <- ggplot(White, aes(x = Depth_disease_transect, y = Max_DHW)) +  scale_y_continuous(name = "Max DHW") + scale_x_discrete(name = "Depth (m)") + geom_boxplot()
p34

#Oxygen
White$Depth_disease_transect<-factor(White$Depth_disease_transect)  #Convert depth into a labelled factor to use as a grouping variabel
p35 <- ggplot(White, aes(x = Depth_disease_transect, y = Oxygen)) + scale_y_continuous(name = "Dissolved Oxygen"~("mg L"^{-1})) + scale_x_discrete(name = "Depth (m)") + 
     geom_boxplot()
p35

#Rainfall
White$Depth_disease_transect<-factor(White$Depth_disease_transect)  #Convert depth into a labelled factor to use as a grouping variabel
p36 <- ggplot(White, aes(x = Depth_disease_transect, y = Rainfall)) + scale_y_continuous(name = "Rainfall (inches)") + scale_x_discrete(name = "Depth (m)") + 
geom_boxplot()
p36

#Salinity
White$Depth_disease_transect<-factor(White$Depth_disease_transect)  #Convert depth into a labelled factor to use as a grouping variabel
p37 <- ggplot(White, aes(x = Depth_disease_transect, y = Salinity)) + scale_y_continuous(name = "Salinity (PSU)") + scale_x_discrete(name = "Depth (m)") + geom_boxplot()
p37

#Temperature
White$Depth_disease_transect<-factor(White$Depth_disease_transect)  #Convert depth into a labelled factor to use as a grouping variabel
p38 <- ggplot(White, aes(x = Depth_disease_transect, y = Temperature)) + scale_y_continuous(name = "Temperature"~(degree*C)) + scale_x_discrete(name = "Depth (m)") + 
     geom_boxplot()
p38

#Turbidity
White$Depth_disease_transect<-factor(White$Depth_disease_transect)  #Convert depth into a labelled factor to use as a grouping variabel
p39 <- ggplot(White, aes(x = Depth_disease_transect, y = Turbidity)) +  scale_y_continuous(name = "Turbidity (NTU)") + scale_x_discrete(name = "Depth (m)") + 
geom_boxplot()
p39

#WP prevalence
White$Depth_disease_transect<-factor(White$Depth_disease_transect)  #Convert depth into a labelled factor to use as a grouping variabel
p40 <- ggplot(White, aes(x = Depth_disease_transect, y = White_Plague_Pervalence)) + scale_y_continuous(name = "WP Prevalence (diseased/total)") + scale_x_discrete(name = "Depth (m)") + geom_boxplot()
p40
```

ENVIRONMENTAL PER REEF TYPE

```{r}
#Chlorophyll
White$Reef.Type<-factor(White$Reef.Type)  #Convert depth into a labelled factor to use as a grouping variabel
p41 <- ggplot(White, aes(x = Reef.Type, y = Chlorophyll))+ scale_y_continuous(name = "Chlorophyll"~("mg m"^{-3})) + scale_x_discrete(name = "Reef Type") +
      geom_boxplot()
p41


#Conductivity
White$Reef.Type<-factor(White$Reef.Type)  #Convert depth into a labelled factor to use as a grouping variabel
p42 <- ggplot(White, aes(x = Reef.Type, y = Conductivity)) +scale_y_continuous(name = "Conductivity"~("COS m"^{-1})) + scale_x_discrete(name = "Reef Type") +
      geom_boxplot()
p42

#Density
White$Reef.Type<-factor(White$Reef.Type)  #Convert depth into a labelled factor to use as a grouping variabel
p43 <- ggplot(White, aes(x = Reef.Type, y = Density)) +scale_y_continuous(name = "Density"~("Kg"%*%"m"^{-3})) + scale_x_discrete(name = "Reef Type") +
        geom_boxplot()
p43

#MAX DHW
White$Reef.Type<-factor(White$Reef.Type)  #Convert depth into a labelled factor to use as a grouping variabel
p44 <- ggplot(White, aes(x = Reef.Type, y = Max_DHW)) +  scale_y_continuous(name = "Max DHW") + scale_x_discrete(name = "Reef Type") + geom_boxplot()
p44

#Oxygen
White$Reef.Type<-factor(White$Reef.Type)  #Convert depth into a labelled factor to use as a grouping variabel
p45 <- ggplot(White, aes(x = Reef.Type, y = Oxygen)) + scale_y_continuous(name = "Dissolved Oxygen"~("mg L"^{-1})) + scale_x_discrete(name = "Reef type") + 
     geom_boxplot()
p45

#Rainfall
White$Reef.Type<-factor(White$Reef.Type)  #Convert depth into a labelled factor to use as a grouping variabel
p46 <- ggplot(White, aes(x = Reef.Type, y = Rainfall)) + scale_y_continuous(name = "Rainfall (inches)") + scale_x_discrete(name = "Reef Type") + 
geom_boxplot()
p46

#Salinity
White$Reef.Type<-factor(White$Reef.Type)  #Convert depth into a labelled factor to use as a grouping variabel
p47 <- ggplot(White, aes(x = Reef.Type, y = Salinity)) + scale_y_continuous(name = "Salinity (PSU)") + scale_x_discrete(name = "Reef Type") + geom_boxplot()
p47

#Temperature
White$Reef.Type<-factor(White$Reef.Type)  #Convert depth into a labelled factor to use as a grouping variabel
p48 <- ggplot(White, aes(x = Reef.Type, y = Temperature)) + scale_y_continuous(name = "Temperature"~(degree*C)) + scale_x_discrete(name = "Reef Type") + 
     geom_boxplot()
p48

#Turbidity
White$Reef.Type<-factor(White$Reef.Type)  #Convert depth into a labelled factor to use as a grouping variabel
p49 <- ggplot(White, aes(x = Reef.Type, y = Turbidity)) +  scale_y_continuous(name = "Turbidity (NTU)") + scale_x_discrete(name = "Reef Type") + 
geom_boxplot()
p49

#WP prevalence
White$Reef.Type<-factor(White$Reef.Type)  #Convert depth into a labelled factor to use as a grouping variabel
p50 <- ggplot(White, aes(x = Reef.Type, y = White_Plague_Pervalence)) + scale_y_continuous(name = "WP Prevalence (diseased/total)") + scale_x_discrete(name = "Reef Type") + geom_boxplot()
p50
```

BOX PLOTS BIOLOGICAL VARIABLES PER YEAR

```{r}
#% Orbicellids
White$Year<-factor(White$Year)  #Convert Month into a labelled factor to use as a grouping variabel
p51 <- ggplot(White, aes(x = Year, y = X.Orbicellids))+ scale_y_continuous(name = "% of Orbicellids") + scale_x_discrete(name = "Years") +
      geom_boxplot()
p51


#% Sand
White$Year<-factor(White$Year)  #Convert Month into a labelled factor to use as a grouping variabel
p52 <- ggplot(White, aes(x = Year, y = X.Sand)) +scale_y_continuous(name = "% of Sand") + scale_x_discrete(name = "Years") +
      geom_boxplot()
p52

#% Coral
White$Year<-factor(White$Year)  #Convert Month into a labelled factor to use as a grouping variabel
p53 <- ggplot(White, aes(x = Year, y = X.Coral)) +scale_y_continuous(name = "% of Coral") + scale_x_discrete(name = "Years") +
        geom_boxplot()
p53

#% Sponges
White$Year<-factor(White$Year)  #Convert Month into a labelled factor to use as a grouping variabel
p54 <- ggplot(White, aes(x = Year, y = X.Sponges)) +  scale_y_continuous(name = "% of Sponges") + scale_x_discrete(name = "Years") + geom_boxplot()
p54

#% Macroalgae
White$Year<-factor(White$Year)  #Convert Month into a labelled factor to use as a grouping variabel
p55 <- ggplot(White, aes(x = Year, y = X.Macroalgae)) + scale_y_continuous(name = "% of Macroalgae") + scale_x_discrete(name = "Years") + 
     geom_boxplot()
p55

#% of Cyanobacteria
White$Year<-factor(White$Year)  #Convert Month into a labelled factor to use as a grouping variabel
p56 <- ggplot(White, aes(x = Year, y = X.Cyanobacteria)) + scale_y_continuous(name = "% of Cyanobacteria") + scale_x_discrete(name = "Years") + 
geom_boxplot()
p56

```

BOX PLOTS BIOLOGICAL VARIABLES PER DEPTH

```{r}
#% Orbicellids
White$Depth_disease_transect<-factor(White$Depth_disease_transect)  #Convert depth into a labelled factor to use as a grouping variabel
p57 <- ggplot(White, aes(x = Depth_disease_transect, y = X.Orbicellids))+ scale_y_continuous(name = "% of Orbicellids") + scale_x_discrete(name = "Depth (m)") +
      geom_boxplot()
p57


#% Sand
White$Depth_disease_transect<-factor(White$Depth_disease_transect)  #Convert depth into a labelled factor to use as a grouping variabel
p58 <- ggplot(White, aes(x = Depth_disease_transect, y = X.Sand)) +scale_y_continuous(name = "% of Sand") + scale_x_discrete(name = "Depth (m)") +
      geom_boxplot()
p58

#% Coral
White$Depth_disease_transect<-factor(White$Depth_disease_transect)  #Convert depth into a labelled factor to use as a grouping variabel
p59 <- ggplot(White, aes(x = Depth_disease_transect, y = X.Coral)) +scale_y_continuous(name = "% of Coral") + scale_x_discrete(name = "Depth (m)") +
        geom_boxplot()
p59

#% Sponges
White$Depth_disease_transect<-factor(White$Depth_disease_transect)  #Convert depth into a labelled factor to use as a grouping variabel
p60 <- ggplot(White, aes(x = Depth_disease_transect, y = X.Sponges)) +  scale_y_continuous(name = "% of Sponges") + scale_x_discrete(name = "Depth (m)") + geom_boxplot()
p60

#% Macroalgae
White$Depth_disease_transect<-factor(White$Depth_disease_transect)  #Convert depth into a labelled factor to use as a grouping variabel
p61 <- ggplot(White, aes(x = Depth_disease_transect, y = X.Macroalgae)) + scale_y_continuous(name = "% of Macroalgae") + scale_x_discrete(name = "Depth (m)") + 
     geom_boxplot()
p61

#% of Cyanobacteria
White$Depth_disease_transect<-factor(White$Depth_disease_transect)  #Convert depth into a labelled factor to use as a grouping variabel
p62 <- ggplot(White, aes(x = Depth_disease_transect, y = X.Cyanobacteria)) + scale_y_continuous(name = "% of Cyanobacteria") + scale_x_discrete(name = "Depth (m)") + 
geom_boxplot()
p62
```

BOX PLOTS BIOLOGICAL VARIABLES PER TYPE OF REEF
```{r}

#% Orbicellids
White$Reef.Type<-factor(White$Reef.Type)  #Convert reef type into a labelled factor to use as a grouping variabel
p63 <- ggplot(White, aes(x = Reef.Type,  y = X.Orbicellids))+ scale_y_continuous(name = "% of Orbicellids") + scale_x_discrete(name = "Reef Type") +
      geom_boxplot()
p63

#% Sand
White$Reef.Type<-factor(White$Reef.Type)  #Convert reef type into a labelled factor to use as a grouping variabel
p64 <- ggplot(White, aes(x = Reef.Type, y = X.Sand)) +scale_y_continuous(name = "% of Sand") + scale_x_discrete(name = "Reef Type (m)") +
      geom_boxplot()
p64

#% Coral
White$Reef.Type<-factor(White$Reef.Type)  #Convert reef type into a labelled factor to use as a grouping variabel
p65 <- ggplot(White, aes(x = Reef.Type, y = X.Coral)) +scale_y_continuous(name = "% of Coral") + scale_x_discrete(name = "Reef Type") +
        geom_boxplot()
p65

#% Sponges
White$Reef.Type<-factor(White$Reef.Type)  #Convert reef type into a labelled factor to use as a grouping variabel
p66 <- ggplot(White, aes(x = Reef.Type, y = X.Sponges)) +  scale_y_continuous(name = "% of Sponges") + scale_x_discrete(name = "Reef Type") + geom_boxplot()
p66

#% Macroalgae
White$Reef.Type<-factor(White$Reef.Type)  #Convert reef type into a labelled factor to use as a grouping variabel
p64 <- ggplot(White, aes(x = Reef.Type, y = X.Macroalgae)) + scale_y_continuous(name = "% of Macroalgae") + scale_x_discrete(name = "Reef Type") + 
     geom_boxplot()
p64

#% of Cyanobacteria
White$Reef.Type<-factor(White$Reef.Type)  #Convert reef type into a labelled factor to use as a grouping variabel
p65 <- ggplot(White, aes(x = Reef.Type, y = X.Cyanobacteria)) + scale_y_continuous(name = "% of Cyanobacteria") + scale_x_discrete(name = "Reef Type") + 
geom_boxplot()
p65

```

BOXPLOTS BIOLOGICAL VARIABLES PER SEASON

```{r}
#BOX PLOTS BIOLOGICAL VARIABLESL PER SEASON
#% Orbicellids
White$Season<-factor(White$Season)  #Convert Month into a labelled factor to use as a grouping variabel
p66 <- ggplot(White, aes(x = Season, y = X.Orbicellids))+ scale_y_continuous(name = "% of Orbicellids") + scale_x_discrete(name = "Seasons") +
      geom_boxplot()
p66


#% Sand
White$Season<-factor(White$Season)  #Convert reef type into a labelled factor to use as a grouping variabel
p67 <- ggplot(White, aes(x = Season, y = X.Sand)) +scale_y_continuous(name = "% of Sand") + scale_x_discrete(name = "Seasons") +
      geom_boxplot()
p67

#% Coral
White$Season<-factor(White$Season)  #Convert reef type into a labelled factor to use as a grouping variabel
p68 <- ggplot(White, aes(x = Season, y = X.Coral)) +scale_y_continuous(name = "% of Coral") + scale_x_discrete(name = "Seasons") +
        geom_boxplot()
p68

#% Sponges
White$Season<-factor(White$Season)  #Convert reef type into a labelled factor to use as a grouping variabel
p69 <- ggplot(White, aes(x = Season, y = X.Sponges)) +  scale_y_continuous(name = "% of Sponges") + scale_x_discrete(name = "Seasons") + geom_boxplot()
p69

#% Macroalgae
White$Season<-factor(White$Season)  #Convert reef type into a labelled factor to use as a grouping variabel
p70 <- ggplot(White, aes(x = Season, y = X.Macroalgae)) + scale_y_continuous(name = "% of Macroalgae") + scale_x_discrete(name = "Seasons") + 
     geom_boxplot()
p70

#% of Cyanobacteria
White$Season<-factor(White$Season)#Convert reef type into a labelled factor to use as a grouping variabel
p71 <- ggplot(White, aes(x = Season, y = X.Cyanobacteria)) + scale_y_continuous(name = "% of Cyanobacteria") + scale_x_discrete(name = "Seasons") + 
geom_boxplot()
p71

```


SIMPLE LINEAR REGRESSION/BINNOMIAL PLOTS - ENVIRONMENTAL VARIABLES VS WHITE PLAGUE PREVALENCE

```{r}
library(ggplot2) 

#using pairs to make scterrplots of specific variables
White1<- White[,c(11:14,28)]
White1
pairs(White1,panel=panel.smooth, line.main= 1, font.labels = 1, col="black",cex.labels = 1.5)


#Chlorophyll vs WP Prevalence
p72 <- ggplot(White, aes(x = Chlorophyll, y = White_Plague_Pervalence)) + scale_x_continuous(name = "Chlorophyll"~("mg m"^{-3})) + scale_y_continuous(name = "WP Prevalence") + geom_point(shape=1) + geom_smooth(method=lm) #adds a linear regression line (95% confidence region)
p72

#Conductivity
p73 <- ggplot(White, aes(y = White_Plague_Pervalence, x = Conductivity)) +scale_x_continuous(name = "Conductivity"~("COS m"^{-1})) + scale_y_continuous(name = "WP Prevalence") + geom_point(shape=1) + geom_smooth(method=lm) #adds a linear regression line (95% confidence region)
p73

#Density
p74 <- ggplot(White, aes(y = White_Plague_Pervalence, x = Density)) +scale_x_continuous(name = "Density"~("Kg"%*%"m"^{-3})) + scale_y_continuous(name = "WP Prevalence") + geom_point(shape=1) + geom_smooth(method=lm) #adds a linear regression line (95% confidence region)
p74

#MAX DHW
p75 <- ggplot(White, aes(y = White_Plague_Pervalence, x = Max_DHW)) +  scale_x_continuous(name = "Max DHW") + scale_y_continuous(name = "WP Prevalence") + geom_point(shape=1) + geom_smooth(method=lm) #adds a linear regression line (95% confidence region)
p75

#Oxygen
p76 <- ggplot(White, aes(y = White_Plague_Pervalence, x = Oxygen)) + scale_x_continuous(name = "Dissolved Oxygen"~("mg L"^{-1})) + scale_y_continuous(name = "WP Prevalence") + geom_point(shape=1) + geom_smooth(method=lm) #adds a linear regression line (95% confidence region)
p76

#Rainfall
p77 <- ggplot(White, aes(y = White_Plague_Pervalence, x = Rainfall)) + scale_x_continuous(name = "Rainfall (mm)") + scale_y_continuous(name = "WP Prevalence") + geom_point(shape=1) + geom_smooth(method=lm) #adds a linear regression line (95% confidence region)
p77

#Salinity
p78 <- ggplot(White, aes(y = White_Plague_Pervalence, x = Salinity)) + scale_x_continuous(name = "Salinity (PSU)") + scale_y_continuous(name = "WP Prevalence") + geom_point(shape=1) + geom_smooth(method=lm) #adds a linear regression line (95% confidence region)
p78

#Temperature
p79 <- ggplot(White, aes(y = White_Plague_Pervalence, x = Temperature)) + scale_x_continuous(name = "Temperature"~(degree*C)) + scale_y_continuous(name = "WP Prevalence") + geom_point(shape=1) + geom_smooth(method=lm) #adds a linear regression line (95% confidence region)
p79

#Turbidity
p80 <- ggplot(White, aes(y = White_Plague_Pervalence, x = Turbidity)) +  scale_x_continuous(name = "Turbidity (NTU)") + scale_y_continuous(name = "WP Prevalence") + geom_point(shape=1) + geom_smooth(method=lm) #adds a linear regression line (95% confidence region)
p80

#Depth
p81 <- ggplot(White, aes(y = White_Plague_Pervalence, x = Depth_disease_transect)) +  scale_x_continuous(name = "Depth (m)") + scale_y_continuous(name = "WP Prevalence") + geom_point(shape=1) + geom_smooth(method=lm) #adds a linear regression line (95% confidence region)
p81

dev.off()
source("http://peterhaschke.com/Code/multiplot.R")
multiplot(p72,p73,p74,p75,p76,p77,p78,p79,p80,p81, cols=3)

#BINOMIAL LINE
#Example from the web
#ggplot(dat, aes(x=mpg, y=vs)) + geom_point() + 
 # stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE)

#Chlorophyll vs WP Prevalence
p82 <- ggplot(White, aes(x = Chlorophyll, y = White_Plague_Pervalence)) + scale_x_continuous(name = "Chlorophyll"~("mg m"^{-3})) + scale_y_continuous(name = "WP Prevalence") + geom_point(shape=1) + geom_smooth(method="glm", method.args=list(family="binomial"), se=FALSE)
p82

#Conductivity
p83 <- ggplot(White, aes(y = White_Plague_Pervalence, x = Conductivity)) +scale_x_continuous(name = "Conductivity"~("COS m"^{-1})) + scale_y_continuous(name = "WP Prevalence") +  geom_point(shape=1) + geom_smooth(method="glm", method.args=list(family="binomial"), se=FALSE)
p83

#Density
p84 <- ggplot(White, aes(y = White_Plague_Pervalence, x = Density)) +scale_x_continuous(name = "Density"~("Kg"%*%"m"^{-3})) + scale_y_continuous(name = "WP Prevalence")+ geom_point(shape=1) + geom_smooth(method="glm", method.args=list(family="binomial"), se=FALSE)
p84

#MAX DHW
p85 <- ggplot(White, aes(y = White_Plague_Pervalence, x = Max_DHW)) +  scale_x_continuous(name = "Max DHW") + scale_y_continuous(name = "WP Prevalence") + geom_point(shape=1) + geom_smooth(method="glm", method.args=list(family="binomial"), se=FALSE)
p85

#Oxygen
p86 <- ggplot(White, aes(y = White_Plague_Pervalence, x = Oxygen)) + scale_x_continuous(name = "Dissolved Oxygen"~("mg L"^{-1})) + scale_y_continuous(name = "WP Prevalence") + geom_point(shape=1) + geom_smooth(method="glm", method.args=list(family="binomial"), se=FALSE)
p86

#Rainfall
p87 <- ggplot(White, aes(y = White_Plague_Pervalence, x = Rainfall)) + scale_x_continuous(name = "Rainfall (mm)") + scale_y_continuous(name = "WP Prevalence") + geom_point(shape=1) + geom_smooth(method="glm", method.args=list(family="binomial"), se=FALSE)
p87

#Salinity
p88 <- ggplot(White, aes(y = White_Plague_Pervalence, x = Salinity)) + scale_x_continuous(name = "Salinity (PSU)") + scale_y_continuous(name = "WP Prevalence") + geom_point(shape=1) + geom_smooth(method="glm", method.args=list(family="binomial"), se=FALSE)
p88

#Temperature
p89 <- ggplot(White, aes(y = White_Plague_Pervalence, x = Temperature)) + scale_x_continuous(name = "Temperature"~(degree*C)) + scale_y_continuous(name = "WP Prevalence") + geom_point(shape=1) + geom_smooth(method="glm", method.args=list(family="binomial"), se=FALSE)
p89

#Turbidity
p90 <- ggplot(White, aes(y = White_Plague_Pervalence, x = Turbidity)) +  scale_x_continuous(name = "Turbidity (NTU)") + scale_y_continuous(name = "WP Prevalence") + geom_point(shape=1) + geom_smooth(method="glm", method.args=list(family="binomial"), se=FALSE)
p90

#Depth
p91 <- ggplot(White, aes(y = White_Plague_Pervalence, x = Depth_disease_transect)) +  scale_x_continuous(name = "Depth (m)") + scale_y_continuous(name = "WP Prevalence") + geom_point(shape=1) + geom_smooth(method="glm", method.args=list(family="binomial"), se=FALSE)
p91

dev.off()
source("http://peterhaschke.com/Code/multiplot.R")
multiplot(p82,p83,p84,p85,p86,p87,p88,p89,p90,p91, cols=3)


```


VARIABLES INTERACTIONS TESTING- Likelihood Ratio test using GLM

```{r}
# Using GLM (family= Binomial) # for White Plague Pervalence in relation to other environmental variables

str(White)

#Evaluating WP prevalence in relation to Depth and Turbidity
White$Depth_disease_transect <-as.factor(White$Depth_disease_transect) # Change Depth as factor
#Give number of corals with and without diseases for the binomial, as it needs both to run
not.ill<- White$ntot-White$ill
not.ill
ill<-White$ill
ill
y<-cbind(ill,not.ill) # binding both vectors of ill and no.ill corals - binomial

m1<- glm(y~White$Depth_disease_transect+White$Turbidity,family= binomial)
print(summary(m1))

m2<-glm(y~White$Depth_disease_transect*White$Turbidity, family=binomial)
print(summary(m2))
anova(m1,m2, test="LRT")


#Evaluating WP prevalence in relation to Turbidity and MAX degree Heating Weeks (MAX DHW)

m3<- glm(y~White$Turbidity+White$Max_DHW,family= binomial)
print(summary(m3)) #significant
m4<-glm(y~White$Turbidity*White$Max_DHW, family= binomial)
print(summary(m4)) # Significant
anova(m3,m4, test="LRT") # Significant



#Check for optimum for Temperature, Oxygen and Salinity?
#optimum :  y = x+x2

m5<- glm(y~(White$Temperature+(White$Temperature)^2)+(White$Oxygen+(White$Oxygen)^2)+(White$Salinity+(White$Salinity)^2),family= binomial)
print(summary(m5)) #significant
m6<-glm(y~(White$Temperature+(White$Temperature)^2)*(White$Oxygen+(White$Oxygen)^2)*(White$Salinity+(White$Salinity)^2), family= binomial)
print(summary(m6)) # Significant
anova(m5,m6, test="LRT") # Significant

#Check for optimum for Temperature, Oxygen and Salinity?
#y=x, 

m9<- glm(y~White$Temperature+White$Oxygen+White$Salinity,family= binomial)
print(summary(m9)) #significant
m10<-glm(y~White$Temperature*White$Oxygen*White$Salinity, family= binomial)
print(summary(m10)) # Significant
anova(m9,m10, test="LRT") # Significant

#Turbidity, Depth and MAx DHW in one model

m7<- glm(y~(White$Turbidity)+(White$Max_DHW)+(White$Depth_disease_transect),family= binomial)
print(summary(m7)) #significant
m8<-glm(y~(White$Turbidity)*(White$Max_DHW)*(White$Depth_disease_transect), family= binomial)
print(summary(m8)) # Significant
anova(m7,m8, test="LRT") # Significant

#Clorophyll, disolved oxygen 

m11<- glm(y~White$Chlorophyll+White$Oxygen,family= binomial)
print(summary(m11)) #significant
m12<-glm(y~White$Chlorophyll*White$Oxygen,family= binomial)
print(summary(m12)) # Significant
anova(m11,m12, test="LRT") # Significant

#Rain and Turbidity

m13<- glm(y~White$Turbidity+White$Rainfall,family= binomial)
print(summary(m13)) #significant
m14<-glm(y~White$Turbidity*White$Rainfall,family= binomial)
print(summary(m14)) # Significant
anova(m13,m14, test="LRT") # Significant


#Only Rain
m15<- glm(y~White$Rainfall,family= binomial)
print(summary(m15)) #significant

#Only Salinityy
m16<-glm(y~White$Salinity,family= binomial)
print(summary(m16)) # NO Significant

#Only Conductivity
m17<-glm(y~White$Conductivity,family= binomial)
print(summary(m17)) #  Significant

#Only Density
m18<-glm(y~White$Density,family= binomial)
print(summary(m18)) # Significant


```


Check data for correlations
```{r}
#Eliminate Month-text column, as we have a Monthnumber column
#White.filtered$Month <- NULL

#Subset the data for environmental variables
Env1 <- data.frame(White[,c(11:13,15:17)]) #First six Environmental variables at a time
Env2<- data.frame(White[,18:21]) # Four more environmental variables
Env3<- data.frame(White[,c(11:13,18:19)]) #Three variables from Env1 and two from Env2
Env4<- data.frame(White[,c(11:13,20:21)]) #Other three variables from Env1 and three from Env2
Env5<- data.frame(White[,c(15:17,18:19)])# Last three env variables fro, Env1 and  two of Env2
Env6<- data.frame(White[,c(15:17,20:21)])# Last three env variables fro, Env1 and  three of Env2

Env7<- data.frame((White[,c(18:19)]))
head(Env2)

#Exploring data summary
#summary(White.filtered)

#use ggpairs to check possible correlations between all 11 environmental variables

library(GGally)
ggpairs(Env1) # Fig Env1. correlated salinity with depth and density with depth
ggpairs(Env2) # Fig En2. No correlation rain and turbidity not clear correlated when using all the stations shallow and deep
ggpairs(Env3) #conductivity and temperature are correlated, and density and temperature

ggpairs(Env4)
ggpairs(Env5)
ggpairs(Env6)
ggpairs(Env7)

pairs(Env1)
pairs(Env2)


cor.test(White$Temperature, White$Temperature)

#Correlated environmental variables include:
White$Depth_disease_transect <-as.numeric(White$Depth_disease_transect) # Change Depth as factor
mean(White$Temperature[White$Depth_disease_transect=="Depth_disease_transect"])
mean(White$Temperature)
White$Depth_disease_transect

#Correlations found:

#temperature- conductivity
#temperature-Density
#Turbidity and Rainfall as not too much correlated when including shallow sites.
#We decided to keep all variables for Bayesian Model. 
```
