---
title: "Environmental drivers of white plague disease on shallow and mesophotic coral reefs, U.S. Virgin Islands- Annual Model Biological and Environmental combined"
author: 
-name:"Andia Chaves Fonnegra and Bernd Panassiti"
  affiliation: University of the Virgin Islands (US.VI) and Laimburg Research Center (Italy)
  abstract: "This document provides the r-code for the white plague disease model described on the article:"Environmental drivers of white plague disease on shallow and mesophotic coral reefs, U.S. Virgin Islands" Description of the model in terms of data collection, white plague disease, parameter estimations and fitting of the model can be found in Appendix A."
  keywords: white plague, coral reefs, bayesian model, environment
date: "7/18/2017"
output: pdf_document
number_sections: yes
    toc: yes
header-includes: \usepackage{graphicx}
email: \email{}
---

```{r setup, include=FALSE, warnings=FALSE}
library(knitr)
opts_knit$set(root.dir='../')                     # definining working directory; or normalizePath('../')
opts_chunk$set(fig.align='center',                # aligns all figures
                echo=TRUE,                        # shows r-code
                message=FALSE,                    # suppresses library outputs
                warnings=FALSE,                   # suppresses library outputs
                tidy=TRUE,  # prevents the source code from running off a pdf page
                dev='pdf')                        # pdf device
```

#Introduction

This document provides the r-code for the white plague disease model described on the article.The model address to the objectives of this study: (1) to evaluate the annual patterns of WP disease, and (2) to identify and quantify the influence of environmental and BIOLOGICAL factors on WP disease prevalence using Bayesian inference. We use rstan for statistical inference (http://mc-stan.org/about/). 

#Data
The white plague prevalence data and corresponding environmental and biological variables used in this model can be downloaded from xxxxxxx. The file "whiteplague_working.RData" needs to be placed in a folder called "Data".
"

#Model

## Load data
```{r}
rm(list=ls(all=TRUE))
# Load settings
source("r-code/00_settings.r")

load(file="data/Whiteplague_workingdata.Rdata")
```


##Load libraries and create directory for results
```{r}
# Date format - forms part of names of created files or graphs
today <- format(Sys.time(), "%Y.%m.%d")
set.seed(2017)

# load libraries
library(formatR)
library(StanHeaders)
library(rstan)
rstan_options(auto_write = TRUE) # automatically save the compilied Stan program
options(mc.cores = parallel::detectCores()) # execute multiple Markov chains in paral

library(MCMCpack) # rwish function
library(modeest)
library(IDPmisc)
library(ggplot2) # for graphs
library(DHARMa) #to extract posterior
```



## Data preparation
```{r}
df <- data.frame(Whiteplague_workingdata_Annual[,c("Ntot","ill","Plague","Chl_sca","Oxy_sca","Sal_sca","Turb_sca","Depth_tran_sca","Temp_sca","DHW_sca", "Sand", "Coral", "Sponges","Macroalgae", "Cyanobacteria")])

<<<<<<< Updated upstream:r-code/03_Whiteplague_rstan_bio_Model.Rmd
df$healthy <- df$Ntot-df$ill
df$Year   <- as.factor(Whiteplague_workingdata_Annual$Year)
df$Site    <- as.factor(as.numeric(Whiteplague_workingdata_Annual$Station_code))
df$Obsid   <- as.factor(1:nrow(df))
=======
#Group data per station, year and season and summarize the max value of each variable
a1<- Whiteplague_workingdata %>% 
  group_by(Station_code, Year, Season) %>% 
  summarise (Coral = max(Coral),Sand = max(Sand), Sponges = max(Sponges), Macroalgae = max(Macroalgae), Cyanobacteria = max(Cyanobacteria),ill = max(ill), Ntot=max(Ntot),Plague = max(Plague),Chl_sca = max(Chl_sca),Oxy_sca = max(Oxy_sca),Sal_sca = max(Sal_sca),Turb_sca = max(Turb_sca),Depth_tran_sca = max(Depth_tran_sca),Temp_sca = max(Temp_sca), DHW_sca = max(DHW_sca) )
a1
head(a1)
>>>>>>> Stashed changes:r-code/WP Biological-Enviornmental Model.Rmd

df$Site_index <- as.integer(df$Site)
df$Year_index=as.integer(df$Year)
df$Obsid_index=as.integer(df$Obsid)

#Environment <- df[,c("Chl_sca","Oxy_sca","Sal_sca","Turb_sca","Depth_tran_sca","Temp_sca","DHW_sca","Sand", "Coral", "Sponges","Macroalgae", "Cyanobacteria")]


df$Chl_sca2 <- df$Chl_sca*df$Chl_sca
df$Oxy_sca2 <- df$Oxy_sca*df$Oxy_sca
df$Sal_sca2 <- df$Sal_sca*df$Sal_sca
df$Turb_sca2 <- df$Turb_sca*df$Turb_sca
df$Depth_tran_sca2 <- df$Depth_tran_sca*df$Depth_tran_sca
df$Temp_sca2 <- df$Temp_sca*df$Temp_sca
df$DHW_sca2 <- df$DHW_sc*df$DHW_sc
df$Sand2 <- df$Sand*df$Sand
df$Coral2 <- df$Coral*df$Coral
df$Sponges2 <- df$Sponges*df$Sponges
df$Macroalgae2 <- df$Macroalgae*df$Macroalgae
df$Cyanobacteria2 <- df$Cyanobacteria*df$Cyanobacteria




# Environment <- df[,c("Chl_sca","Chl_sca2","Oxy_sca","Oxy_sca2","Sal_sca","Sal_sca2","Turb_sca","Turb_sca2","Depth_tran_sca","Depth_tran_sca2","Temp_sca","Temp_sca2",
#                      "DHW_sca","DHW_sca2","Sand", "Sand2","Coral","Coral2", "Sponges","Sponges2","Macroalgae", "Macroalgae2","Cyanobacteria","Cyanobacteria2")]


# Environment <- df[,c("Chl_sca","Chl_sca2","Oxy_sca","Oxy_sca2","Sal_sca","Sal_sca2","Turb_sca","Turb_sca2","Depth_tran_sca","Depth_tran_sca2","Temp_sca","Temp_sca2", "DHW_sca","DHW_sca2",
#                      "Sand","Coral","Sponges","Macroalgae", "Cyanobacteria")]


Environment <- df[,c("Chl_sca","Chl_sca2","Oxy_sca","Oxy_sca2","Sal_sca","Sal_sca2","Turb_sca","Depth_tran_sca","Temp_sca","Temp_sca2", "DHW_sca","DHW_sca2",
                     "Sand","Coral","Sponges","Macroalgae", "Cyanobacteria")]



# df$Year1 <- rep(0,length(df$Year_index))
# df$Year2 <- rep(0,length(df$Year_index))
# df$Year3 <- rep(0,length(df$Year_index))
# 
# df$Year1[which(df$Year_index==1)]<-1
# df$Year2[which(df$Year_index==2)]<-1
# df$Year3[which(df$Year_index==3)]<-1

#Environment <- df[,c("Chl_sca","Oxy_sca","Sal_sca","Turb_sca","Depth_tran_sca","Temp_sca","DHW_sca","Sand", "Coral", "Sponges","Macroalgae", "Cyanobacteria","Year1","Year2","Year3")]
```


## Model definition
```{r}
dataList = list(
  TotalObservedcolonies = df$Ntot, #Total observed colonies Shallow (2 years) Deep (4 years)
  Diseasedcolonies = df$ill, # Total numer of disease colonies Shallow (2 years) Deep (4 years)
  nWP = nrow(df), #White Plague prevalence total number of rows
  observedWpPrevalence = df$Plague, # Total Prevalence of White Plague Shallow (2 years) Deep (4 years)
  
  Environment = Environment,
  nEnvi = ncol(Environment),
  
  Year = df$Year_index,
  nYear =length(unique(df$Year_index))
)
```


## Running the model

```{r model_run,eval=FALSE}
time <- proc.time()
stan.white <- stan("r-code/03_Whiteplague_rstan_bio_Model.stan",
                   data = dataList,
                   iter = 300, 
                   chains = 4,
                   control = list(adapt_delta = 0.99,max_treedepth = 15),
                   refresh=10)
(proc.time() - time)/60/60
```

```{r}
NvarEnvironment <- ncol(Environment)
printVarEnvironment <- paste(rep("parEnvironment[",NvarEnvironment),1:NvarEnvironment,rep("]",NvarEnvironment),sep="")

print(stan.white,pars=c("b0",printVarEnvironment),probs=c(.1,.5,.9))
plot(stan.white,pars=c("b0",printVarEnvironment))


#plotting the posterior distribution for the parameters
post_beta<-As.mcmc.list(stan.white,pars=c("b0",printVarEnvironment))
plot(post_beta)
#traceplot(post_beta, inc_warmup = FALSE)
```




```{r}
posterior<-extract(stan.white,inc_warmup=F,permute=F)

Ntot <- Whiteplague_workingdata_Annual$Ntot
observedWhitePlague <- Whiteplague_workingdata_Annual$ill

Nchains=4 # number of chains, see model definition
modelpredictionsChains <- paste(rep("modelPredictions[,,",Nchains),1:Nchains,rep("]",Nchains),sep="")


# FITTED PREDICTED
modelPredictions <- NULL
modelPredictions<-apply(posterior, MARGIN=1:2, FUN = function(draw) draw[grepl("^RealDisease[\\[]",names(draw))])


# combine 4 chains
predictionsLinearWhitePlague <- NULL
assign("predictionsLinearWhitePlague",eval(parse(text=paste("cbind(",paste(modelpredictionsChains, collapse = ","),")",sep=""))))

predictionsResponseMedianWhitePlague <- inv.logit(as.vector(apply(predictionsLinearWhitePlague, 1, median)))*Ntot

# SIMULATED
modelPredictions <- NULL
modelPredictions<-apply(posterior, MARGIN=1:2, FUN = function(draw) draw[grepl("^DiseasePredictions[\\[]",names(draw))])

# combine 4 chains
simulatedPredictedWhitePlague <- NULL
assign("simulatedPredictedWhitePlague",eval(parse(text=paste("cbind(",paste(modelpredictionsChains, collapse = ","),")",sep=""))))


WhitePlagueDHARMaResponse <- createDHARMa(simulatedResponse = simulatedPredictedWhitePlague,
  observedResponse = observedWhitePlague,fittedPredictedResponse = predictionsResponseMedianWhitePlague,integerResponse = T)


plotSimulatedResiduals(simulationOutput = WhitePlagueDHARMaResponse)
```



#Save the model run results
```{r}
save(stan.white, file =paste("data/2018.02.26a_Whiteplague_model_result.Rdata",sep=""))
```




###Test overdispersion
```{r}
#testOverdispersion(simulationOutput = WhitePlagueDHARMaResponse)
```
###Test zero inflation
```{r}
#testZeroInflation(WhitePlagueDHARMaResponse)
```

#Save the model run results
```{r}
save(stan.white, file =paste("data/2017.11.21a_Whiteplague_model_result.Rdata",sep=""))
```

# Session information for model run
```{r}
devtools::session_info()
```

#Plotting the model
```{r ploting from stan to ggmcmc}

print(stan.white)



plot(stan.white)

str(stan.white)

library(ggmcmc)
library(coda)

#This is to Install Stan Caterpillar StanCat to build caterpillar graphs from Stan
#devtools::install_github('christophergandrud/StanCat')
StanCat::stan_caterpillar(stan.white, pars = 'v_raw\\[.*\\]')

#To use ggmcmc#the rstan::extract, specify that is extracted from rstan procedure and not from dpylr
library(plyr)
#this one extracts everything but is not plotting histograms properly
s <- rstan::extract(stan.white, inc_warmup=FALSE, permuted=FALSE) 
#this one extracts only some functions

s <- mcmc.list(mcmc(s[,1,-4]), mcmc(s[,2,-4]), mcmc(s[,3,-4]), mcmc(s[,4,-4]))
S <- ggs(s) 
ggmcmc(S)
ggs_traceplots(S)
ggs_caterpillar(S)


#for specific parameters. But is not runing
#s <- rstan::extract(stan.white, pars = 'v_raw', permuted = TRUE)$parEnvironment 
#s <- do.call(mcmc.list, alply(s[,,-4], 2, mcmc))

```

#Check model output with Shiny
```{r extract with shiny}
library(shiny)
library(shinystan)

launch_shinystan(stan.white)
```


#Save r markdown file
```{r-To save the r markdown file}

save(dataList, stan.white, file = paste("Figures/", today, "_fit.fullModel.pdf",
sep = ""))
save(dataList, file = paste("Figures/", today, "_fit.fullModel_data.Rdata",
sep = ""))
library(lattice)
trellis.device("pdf", file = paste("Figures/", today, "_Output_rstan.pdf",
sep = ""), width = 14, height = 7)
plot(stan.white)
dev.off()


library(lattice)
trellis.device("pdf")
file= paste("Figures/",today,"_Output_rstan.pdf",sep="2-2016", width=14, height=7)
plot(stan.white)
```

##Cauchy and normal distribution test

```{r}
#Normal distribution
x <- seq(-10, 10, length=100)
hx <- dnorm(x)

degf <- c(1, 3, 8, 30)
colors <- c("red", "blue", "darkgreen", "gold", "black")
labels <- c("df=1", "df=3", "df=8", "df=30", "normal")

plot(x, hx, type="l", lty=2, xlab="x value",
  ylab="Density", main="Comparison of t Distributions")

for (i in 1:4){
  lines(x, dt(x,degf[i]), lwd=2, col=colors[i])
}

legend("topright", inset=.05, title="Distributions",
  labels, lwd=2, lty=c(1, 1, 1, 1, 2), col=colors)

```

#cauchy

```{r}
#Cauchy distribution

x <- seq(-10, 10, length=100)
hi <- dcauchy(x)

degf <- c(1, 3, 8, 30)
colors <- c("red", "blue", "darkgreen", "gold", "black")
labels <- c("df=1", "df=3", "df=8", "df=30", "cauchy")

plot(x, hx, type="l", lty=2, xlab="x value",
  ylab="Density", main="Comparison of t Distributions")

for (i in 1:4){
  lines(x, dt(x,degf[i]), lwd=2, col=colors[i])
}

legend("topright", inset=.05, title="Distributions",
  labels, lwd=2, lty=c(1, 1, 1, 1, 2), col=colors)
```

