---
output: html_document
editor_options: 
  chunk_output_type: console
---
---*
title: "Environmental drivers of white plague disease on shallow and mesophotic coral reefs, U.S. Virgin Islands"
subtitel: "Residual diagnosis"
author: "Andia Chaves-Fonnegra, Bernd Panassiti, Tyler B. Smith, Elizabeth Brown, Elizabeth Clemens, Moriah Sevier, and Marilyn Brandt"
date: 'created: 25/07/2017, last modified: `r format(Sys.Date(), format="%d.%m.%Y")`'
output: pdf_document
---

```{r setup, include=FALSE, warnings=FALSE}
library(knitr)
opts_knit$set(root.dir='../')                     # definining working directory; or normalizePath('../')
opts_chunk$set(fig.align='center',                # aligns all figures
                echo=TRUE,                        # shows r-code
                message=FALSE,                    # suppresses library outputs
                warnings=FALSE,                   # suppresses library outputs
                tidy=TRUE,  # prevents the source code from running off a pdf page
                dev='pdf')                        # pdf device
```


# Environmental drivers of white plague disease on shallow and mesophotic coral reefs, U.S. Virgin Islands - Residual analysis

## Introduction
This document provides residual diagnosis for the logistic regression model using Bayesian inference for the White Plague disease on corals described in the article: ``Identifying local drivers of a vector-pathogen-disease system using Bayesian modeling''.
The Skript is based on the residuals analysis provided in Panassiti et al., 2016.

## Data






## Residual analysis


#### Loading data & libraries
```{r}
rm(list=ls(all=TRUE))
#NameOfModelRun <- "2018.03.08_Whiteplague_Environmental_model_result_graphinput" # last environmental
NameOfModelRun <- "2021.01.17_Whiteplague_Biological_model_result_graphinput" # last biological
load(file=paste("results/",NameOfModelRun,".Rdata",sep=""))
load(file="data/Whiteplague_workingdata.Rdata")

#"devtools::install_github(repo = "DHARMa", username = "florianhartig", subdir = "DHARMa", dependencies = T, build_vignettes = TRUE)

library(DHARMa)
library(ggplot2)
library(lattice)
```


```{r}
# Biological model
ill <- Whiteplague_workingdata_Annual$ill
Latitude <- Whiteplague_workingdata_Annual$Latitude
Longitude <- Whiteplague_workingdata_Annual$Longitude
Year <- Whiteplague_workingdata_Annual$Year
#Month_12 <- Whiteplague_workingdata_Annual$Month_12 # NOT READY
Station_code <- Whiteplague_workingdata_Annual$Station_code

# Environmental model
# ill <- Whiteplague_workingdata$ill
# Latitude <- Whiteplague_workingdata$Latitude
# Longitude <- Whiteplague_workingdata$Longitude
# Year <- Whiteplague_workingdata$Year
# Month_12 <- Whiteplague_workingdata$Month_12
# Station_code <- Whiteplague_workingdata$Station_code
```


#### Create DHARMa object
```{r,echo=FALSE}
# Symptomatic corals - Response scale - inv.logit
set.seed(2021)
WhitePlagueDHARMaResponse <- createDHARMa(simulatedResponse = simulatedPredictedWhitePlague,
  observedResponse = ill,fittedPredictedResponse = predictionsResponseMedianWhitePlague,integerResponse = T)
```


#### Plot DHARMa object
```{r}
plot(WhitePlagueDHARMaResponse)
```


#### Spatial autocorrelation
```{r}
#test Spatial autocorrelation Moran's I

# Problem repeating sites result in 0 distances!
library(dplyr)
df <- data.frame(scaledResiduals=WhitePlagueDHARMaResponse$scaledResiduals,x=Whiteplague_workingdata_Annual$Latitude,y=Whiteplague_workingdata_Annual$Longitude)

# 1 option - group and mean

dfNew <- data.frame(df %>%
  group_by(x, y) %>%
  summarize(scaledResiduals = mean(scaledResiduals, na.rm = TRUE)))

testSpatialAutocorrelation(simulationOutput = dfNew$scaledResiduals, x = dfNew$x,y = dfNew$y)

# 	DHARMa Moran's I test for spatial autocorrelation
# 
# data:  dfNew$scaledResiduals
# observed = -0.307581, expected = -0.083333, sd = 0.165954, p-value = 0.1766
# alternative hypothesis: Spatial autocorrelation
# 
# 
# 2. option - random noise
df$x_new <- rnorm(df$x,df$x,0.001)
df$y_new <- rnorm(df$y,df$y,0.001)

testSpatialAutocorrelation(simulationOutput = df$scaledResiduals, x = df$x_new,y = df$y_new)
```

Both tests indicate that we should reject the null of no spatial auto-correlation. There is spatial autocorrelation.

### Test for residuals temporal autocorrelation
```{r}
df <- data.frame(scaledResiduals=WhitePlagueDHARMaResponse$scaledResiduals,x=Whiteplague_workingdata_Annual$Latitude,y=Whiteplague_workingdata_Annual$Longitude)

# test temporal autocorrelation Durbin-Watson test in the residuals


#Year
dfNew=recalculateResiduals(WhitePlagueDHARMaResponse, group = Whiteplague_workingdata_Annual$Year)
testTemporalAutocorrelation(simulationOutput = dfNew, time=1:3,plot=T)

#Random time
testTemporalAutocorrelation(simulationOutput = WhitePlagueDHARMaResponse, time = "random")
```



#Predictive maps of disease
#Figure 5. Standardized residuals (Bayesian p-values) for the two models of WP.


```{r}
library(maptools)
#crswgs84=CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")


library(rgdal)     # R wrapper around GDAL/OGR
library(ggplot2)   # for general plotting
library(ggmap)    # for fortifying shapefiles

# First read in the shapefile, using the path to the shapefile and the shapefile name minus the
# extension as arguments
shapefile <- readOGR("data/GIS/", "tl_2015_78_estate_BP")
SHP_study_area <- shapefile

# Next the shapefile has to be converted to a dataframe for use in ggplot2
#SHP_study_area <- fortify(shapefile)


# Now the shapefile can be plotted as either a geom_path or a geom_polygon.
# Paths handle clipping better. Polygons can be filled.
# You need the aesthetics long, lat, and group.
# map <- ggplot() +
# geom_path(data = SHP_study_area, 
#           aes(x = long, y = lat, group = group),
#           color = 'gray', size = .2)+
# geom_point(aes(x = Whiteplague_workingdata_Annual$Longitude, y = Whiteplague_workingdata_Annual$Latitude) ,shape=23, fill="blue", color="darkred", size=3)
# #print(map)
# # Using the ggplot2 function coord_map will make things look better and it will also let you change
# # the projection. But sometimes with large shapefiles it makes everything blow up.
# map_projected <- map +
#   coord_map()
#   
# print(map_projected)
```



##### Observed incidenes per site
```{r}
library("ggplot2")
library("maps")
library("grid")
library("gridExtra")


colFunc<-colorRampPalette(c("blue","white","red"))
colFuncBlueRed<-colFunc(10)


x = Longitude
y = Latitude
df <- data.frame(x,y,values=WhitePlagueDHARMaResponse$scaledResiduals,sites=Station_code)

SHP_study_area_Fortified <- fortify(shapefile)


p1 <- ggplot() +
geom_path(data = SHP_study_area_Fortified, aes(x = long, y = lat, group = group),color = 'gray', size = .2)+
#geom_point(data=df,aes(x = x, y = y) ,shape=23, fill="blue", color="darkred", size=3)+
#geom_boxplot(data=df, aes(x=x, y=y, stat = "identity",fill=factor(sites), group=sites))+
  theme_bw()+
coord_cartesian(xlim=c(-65.2,-64.5),ylim=c(18.14,18.42))

```


```{r}
grid.newpage()
pushViewport(dataViewport(x, y, c(-65.2,-64.5), c(18.14,18.42)))
#pushViewport(dataViewport(x, y, c(0,1), c(0,1)))
v1<-viewport(default.units = "native") #plot area for the main map

print(p1,vp=v1) 

#scales::show_col(colFuncBlueRed)

for (i in 1:length(unique(df$sites))){
  dfSite <- df[df$sites %in% unique(df$sites)[i],]
  

  x<-mean(dfSite$values)
  if(x <=0.1){pointColor<-colFuncBlueRed[1]}
  if(x > 0.1 & x<=0.2){pointColor<-colFuncBlueRed[2]}
  if(x > 0.2 & x<=0.3){pointColor<-colFuncBlueRed[3]}
  if(x > 0.3 & x<=0.4){pointColor<-colFuncBlueRed[4]}
  if(x > 0.4 & x<=0.5){pointColor<-colFuncBlueRed[5]}
  if(x > 0.5 & x<=0.6){pointColor<-colFuncBlueRed[6]}
  if(x > 0.6 & x<=0.7){pointColor<-colFuncBlueRed[7]}
  if(x > 0.7 & x<=0.8){pointColor<-colFuncBlueRed[8]}
  if(x > 0.8 & x<=0.9){pointColor<-colFuncBlueRed[9]}
  if(x > 0.9 & x<=1  ){pointColor<-colFuncBlueRed[10]}
  

  
p2 <- ggplot(dfSite, aes(sites, values)) + geom_boxplot(fill=pointColor)+ 
  theme(
      axis.title=element_blank()
    ,  panel.background = element_rect(fill = "transparent") # bg of the panel
    , plot.background = element_rect(fill = "transparent") # bg of the plot
    , panel.grid.major = element_blank() # get rid of major grid
    , panel.grid.minor = element_blank() # get rid of minor grid
    , legend.background = element_rect(fill = "transparent") # get rid of legend bg
    , legend.box.background = element_rect(fill = "transparent") # get rid of legend panel bg
  )

  if (i==1){dfSite$x[1] <- -64.98}
v2<-viewport(width = 0.05, height = 0.05, x = dfSite$x[1], y = dfSite$y[1],default.units = "native") #plot area for the inset map

# 
# x=as.numeric(convertX( unit(dfSite$x[1],'native'),'npc'))
# 
# x=convertX( unit(dfSite$x[1],'native'),'npc',valueOnly=TRUE)
# valueOnly
# 
# y=as.numeric(convertX( unit(dfSite$y[1],'native'),'in'))
# v2<-viewport(width = 0.5, height = 0.5, x = x, y = y)

print(p2,vp=v2)

}
```

