---
title: "Environmental drivers of white plague disease on shallow and mesophotic coral reefs, U.S. Virgin Islands"
subtitel: "Residual diagnosis"
author: "Andia Chaves-Fonnegra, Bernd Panassiti, Tyler B. Smith, Elizabeth Brown, Elizabeth Clemens, Moriah Sevier, and Marilyn Brandt"
date: 'created: 25/07/2017, last modified: `r format(Sys.Date(), format="%d.%m.%Y")`'
output: pdf_document
---

```{r setup, include=FALSE, warnings=FALSE}
library(knitr)
opts_knit$set(root.dir='../')                     # definining working directory; or normalizePath('../')
opts_chunk$set(fig.align='center',                # aligns all figures
                echo=TRUE,                        # shows r-code
                message=FALSE,                    # suppresses library outputs
                warnings=FALSE,                   # suppresses library outputs
                tidy=TRUE,  # prevents the source code from running off a pdf page
                dev='pdf')                        # pdf device
```


# Environmental drivers of white plague disease on shallow and mesophotic coral reefs, U.S. Virgin Islands - Residual analysis

## Introduction
This document provides residual diagnosis for the logistic regression model using Bayesian inference for the White Plague disease on corals described in the article: ``Identifying local drivers of a vector-pathogen-disease system using Bayesian modeling''.
The Skript is based on the residuals analysis provided in Panassiti et al., 2016.

## Data






## Residual analysis


#### Loading data & libraries
```{r}
rm(list=ls(all=TRUE))
<<<<<<< HEAD

load(file="data/Whiteplague_workingdata.RData)

NameOfModelRun <- "2017.07.26_Whiteplague_model_result_graphinput"
load(file=paste("data/",NameOfModelRun,".RData",sep=""))
=======
load(file="data/Whiteplague_workingdata.Rdata")

NameOfModelRun <- "2017.07.26_Whiteplague_model_result"
load(file=paste("data/",NameOfModelRun,"_graphinput.Rdata",sep=""))
>>>>>>> origin/master

#devtools::install_github(repo = "DHARMa", username = "florianhartig", subdir = "DHARMa", dependencies = T, build_vignettes = TRUE)
library(DHARMa)
```


#### Create DHARMa object
```{r,echo=FALSE}
# Symptomatic corals - Response scale - inv.logit
WhitePlagueDHARMaResponse <- createDHARMa(simulatedResponse = simulatedPredictedWhitePlague,
  observedResponse = observedWhitePlague,fittedPredictedResponse = predictionsResponseMedianWhitePlague,integerResponse = T)
```


#### Plot DHARMa object
```{r}
plotSimulatedResiduals(simulationOutput = WhitePlagueDHARMaResponse)
```

#### Spatial autocorrelation
```{r}
#test Spatial autocorrelation Moran's I

# Problem repeating sites result in 0 distances!
library(dplyr)
df <- data.frame(scaledResiduals=WhitePlagueDHARMaResponse$scaledResiduals,x=Whiteplague_workingdata$Latitude,y=Whiteplague_workingdata$Longitude)

# 1 option - group and mean

dfNew <- as.data.frame(df %>%
  group_by(x, y) %>%
  summarize(scaledResiduals = mean(scaledResiduals, na.rm = TRUE)))
testSpatialAutocorrelation(simulationOutput = dfNew, x = dfNew$x,y = dfNew$y)

# 2. option - random noise
df$x_new <- rnorm(df$x,df$x,0.001)
df$y_new <- rnorm(df$y,df$y,0.001)

testSpatialAutocorrelation(simulationOutput = df, x = df$x_new,y = df$y_new)
```

Both tests indicate that we should reject the null of no spatial auto-correlation.