---
title: "Whiteplague_rstanModel"
author: "Andia Chaves Fonnegra"
date: "7/18/2017"
output: pdf_document
---

```{r setup, include=FALSE, warnings=FALSE}
library(knitr)
opts_knit$set(root.dir='../')                     # definining working directory; or normalizePath('../')
opts_chunk$set(fig.align='center',                # aligns all figures
                echo=TRUE,                        # shows r-code
                message=FALSE,                    # suppresses library outputs
                warnings=FALSE,                   # suppresses library outputs
                tidy=TRUE,  # prevents the source code from running off a pdf page
                dev='pdf')                        # pdf device
```



Install Packages to run rstan in R if necessary
```{r Install apckages if required}

#Install packages and load libraries to run Stan and graphing results from Stan
if(!require(formatR))
{
  install.packages("formatR", repos = mir)
  require(formatR)
}

if(!require(StanHeaders)) 
{
  install.packages("StanHeaders", repos = mir)
  require(StanHeaders)
}

if(!require(rstan))
{
  install.packages("rstan", repos = mir)
  require(rstan)
}

if(!require(MCMCpack)) # rwish function
{
  install.packages("MCMCpack", repos = mir)
  require(MCMCpack)
}

if(!require(modeest)) 
{
  install.packages("modeest", repos = mir)
  require(modeest)
} 

if(!require(IDPmisc))
{
  install.packages("IDPmisc", repos = mir)
  require(IDPmisc)
} 

if(!require(ggplot2)) 
{
  install.packages("ggplot2", repos = mir)
  require(ggplot2)
}
if(!require(shinystan))
{
  install.packages("shinystan", repos = mir)
  require(shinystan)
}
if(!require(qtlcharts))
{
  install.packages("qtlcharts", repos = mir)
  require(qtlcharts)
}
```

Load libraries if required

```{r}
# load libraries
library(formatR)
library(StanHeaders)
library(rstan)
library(MCMCpack) # rwish function
library(modeest)
library(IDPmisc)
library(ggplot2)
# create directory if not existing
suppressWarnings(dir.create(file.path(getwd(),"Results/04_Rstan")))

```

Loading Whiteplague_workeddata.rdata 
```{r}
rm(list=ls(all=TRUE))
# Load settings
#White1<- load('~/Documents/Project-White/Data/Whiteplague_sorted_data.rdata')
#ls()

#check data
ls(White1)
dim(White1)
head(White1)
##grep("^White$Chl_sca", colnames(x))

##Selecting Environmental variables alrready SCALED to use in the model
enviroselection = White1[c(32:42)] # to select all 11 environmental variables. ``{r}
enviroselection

#count how many months we have in total
library(plyr)
count(White1$Month_12)
```



# Data as a list, using Month as factor
```{r}
Month=rep(1:12,c(3,25,26,29,22,4,7,29,30,14,25,25))  # Total of 12 months, 
dataList = list(
        TotalObservedcolonies = White1$Ntot,
        Diseasedcolonies = White1$ill,
        nWP = length(White1$Plague),
        
                  Chlorophyll = White1$Chl_sca,
                  Conductivity = White1$Cond_sca,
                  Density = White1$Den_sca,
                  CTD_Depth = White1$DepCTD_sca,    
                  Oxygen = White1$Oxy_sca,   
                  Salinity = White1$Sal_sca,    
                  Turbidity = White1$Turb_sca,    
                  Depth = White1$Depth_tran_sca,  
                  Temperature = White1$Temp_sca,   
                  Max_DHW = White1$DHW_sca  , 
                  Rainfall = White1$Rain_sca,   
        
        Environment = White1[32:(32-1+length(enviroselection))],
        nEnvi = length(enviroselection),

        Month = Month,
        nMonth = length(unique(Month))
)
```


## Model definition

```{r define_model}
# Define model

model.bn.Wplague <- '
data {
// counters
int<lower=0> nWP;    //number of White Plague cases, int=integer  and non negative >0 
int<lower=0> nEnvi;  //number of environmental variables
int<lower=0> nMonth; //number of Months sampled


// predictors
matrix[nWP, nEnvi] Environment;
int<lower=0> Month[nWP];

// response
int<lower=0> TotalObservedcolonies [nWP]; // Total number of coral colonies (cases)
int<lower=0> Diseasedcolonies [nWP]; // number of diseased coral colonies (successes)
}

parameters {
// parameters that will be used in the model should be defined here
// these will be saved and printed with the output

real b0;                        // intercept
vector[nEnvi] parEnvironment;   //Environmental-level


real <lower=0> tauMonth;
vector[nMonth] v_raw;
}

transformed parameters {
// generation of derived parameters from parameters above should be defined here
// these will be saved and printed with the output


vector[nMonth] v; //month effects
real RealDisease[nWP];
vector[nWP] mu; //


for (h in 1:nMonth){
v[h] = v_raw[h]*(tauMonth);
}

for( i in 1:nWP ) {
mu[i] = v[Month[i]];
RealDisease[i]  = (b0  + Environment[i] * parEnvironment + mu[i]);
}
}


model{
// This is area for describing the model including the priors and likelihood
// This is the only place where sampling can be done within Stan
// All data and parameters used in this section must be defined above

// Fixed effects priors
b0 ~ normal(0, 100 ); //
parEnvironment ~ normal(0 , 100 );

//Hyperpriors
tauMonth  ~ inv_gamma(0.001, 0.001);    

//Random effects priors
v_raw ~ normal(0, 1);

Diseasedcolonies ~ binomial_logit(TotalObservedcolonies,RealDisease);
}


#predictive inference
#generated quantities
#{
#real DiseasePredictions[nWP]; //vector to store predictions

#for( i in 1:nWP ) {
#DiseasePredictions [i] <- binomial_rng(TotalObservedcolonies[i],inv_logit(b0  + Environment[i] * parEnvironment + mu[i]));
#}
#}
'
```

## Running the model
```{r test_run,echo=FALSE,eval=FALSE}
stan.white <- stan(model_code = model.bn.Wplague, data = dataList, iter = 1000, chains = 1,refresh=1,thin=1)
```

```{r model_run,eval=FALSE}
time <- proc.time()
set.seed(2017)
stan.white2 <- stan(model_code = model.bn.Wplague, data = dataList, iter = 30000, chains = 4,refresh=1)
(proc.time() - time)/60/60
```


# Session info
```{r}
devtools::session_info()
```


```{r ploting from stan to ggmcmc}
save(stan.white2, file ="Data/stan.white2")
print(stan.white2)
plot(stan.white2)

str(stan.white2)

library(ggmcmc)
library(coda)

#This is to Install Stan Caterpillar StanCat to build caterpillar graphs from Stan
#devtools::install_github('christophergandrud/StanCat')
StanCat::stan_caterpillar(stan.white2, pars = 'v_raw\\[.*\\]')

#To use ggmcmc#the rstan::extract, specify that is extracted from rstan procedure and not from dpylr
library(plyr)
#this one extracts everything but is not plotting histograms properly
s <- rstan::extract(stan.white2, inc_warmup=FALSE, permuted=FALSE) 
#this one extracts only some functions

s <- mcmc.list(mcmc(s[,1,-4]), mcmc(s[,2,-4]), mcmc(s[,3,-4]), mcmc(s[,4,-4]))
S <- ggs(s) 
ggmcmc(S)
ggs_traceplots(S)
ggs_caterpillar(S)


#for specific parameters. But is not runing
#s <- rstan::extract(stan.white, pars = 'v_raw', permuted = TRUE)$parEnvironment 
#s <- do.call(mcmc.list, alply(s[,,-4], 2, mcmc))

```

```{r-To save the r markdown file}

save(dataList, stan.white, file = paste("Figures/", today, "_fit.fullModel.pdf",
sep = ""))
save(dataList, file = paste("Figures/", today, "_fit.fullModel_data.Rdata",
sep = ""))
library(lattice)
trellis.device("pdf", file = paste("Figures/", today, "_Output_rstan.pdf",
sep = ""), width = 14, height = 7)
plot(stan.white)
dev.off()


library(lattice)
trellis.device("pdf")
file= paste("Figures/",today,"_Output_rstan.pdf",sep="2-2016", width=14, height=7)
plot(stan.white)
```

```{r extract with shiny}
library(shiny)
library(shinystan)
launch_shinystan(stan.white2)
```


#From Panassiti et al., 2016
#Extraction of residuals from the posterior of a stan-object
#At first rstan package is loaded and the entire posterior with all parameters extracted

```{r}
library(rstan)
posterior<-extract(stan.white2,inc_warmup=F,permute=F) # extraction of posterior
```

Extraction of posterior predictive simulations.
```{r}
# real White Plague Prevalence Predictions are simulated (predictions generated in the
# Generated quantities block of the model code)
modelPredictions<-apply(posterior, MARGIN=1:2, FUN = function(draw)
draw[grepl("^RealDisease[\\[]",names(draw))])

# combine 4 chains
simulatedPredictedWP <- NULL
simulatedPredictedWP <- cbind(modelPredictions[,,1], modelPredictions[,,2],
modelPredictions[,,3], modelPredictions[,,4]) # combining chains
```


Extraction of median model predictions on the response scale
```{r}
#library(LaplacesDemon)
#library(arm) 
#library(boot)

# Real WP prevalence Predictions (are the predictions on the linear scale which are
# transformed in a 2. step to be on the response scale (inv.logit()))
modelPredictions <- NULL
modelPredictions<-apply(posterior, MARGIN=1:2, FUN = function(draw)
draw[grepl("^RealDisease[\\[]",names(draw))])
# combine 4 chains
predictionsLinearWP <- NULL
predictionsLinearWP <- cbind(modelPredictions[,,1], modelPredictions[,,2],
modelPredictions[,,3], modelPredictions[,,4]) # combining chains
predictionsResponseWPMedian <-inv.logit(as.vector(apply(predictionsLinearWP, 1, median)))*White1$Plague # I think is WP prevalence and not ill here for White1$Plague..mmm? This one worked
```

Residual diagnosis
```{r}
#devtools::install_github(repo = "DHARMa", username = "florianhartig",
#subdir = "DHARMa", dependencies = T, build_vignettes = TRUE)
library(DHARMa)


WP_DHARMaResponse <- createDHARMa(
simulatedResponse = simulatedPredictedWP,
observedResponse = White1$Plague,
fittedPredictedResponse = predictionsResponseWPMedian,
integerResponse = T) # interger correction function

#Plots
plotSimulatedResiduals(simulationOutput = WP_DHARMaResponse)

#test Spatial autocorrelation Moran's I
testSpatialAutocorrelation(WP_DHARMaResponse , x = trapCoords$st_x,
y = trapCoords$st_y)

```

