---
title: "Environmental drivers of white plague disease on shallow and mesophotic coral reefs, U.S. Virgin Islands- Annual Model Biological and Environmental combined"
author: 
-name:"Andia Chaves Fonnegra and Bernd Panassiti"
  affiliation: University of the Virgin Islands (US.VI) and Laimburg Research Center (Italy)
  abstract: "This document provides the r-code for the white plague disease model described on the article:"Environmental drivers of white plague disease on shallow and mesophotic coral reefs, U.S. Virgin Islands" Description of the model in terms of data collection, white plague disease, parameter estimations and fitting of the model can be found in Appendix A."
  keywords: white plague, coral reefs, bayesian model, environment
date: "7/18/2017"
output: pdf_document
number_sections: yes
    toc: yes
header-includes: \usepackage{graphicx}
email: \email{}
---

```{r setup, include=FALSE, warnings=FALSE}
library(knitr)
opts_knit$set(root.dir='../')                     # definining working directory; or normalizePath('../')
opts_chunk$set(fig.align='center',                # aligns all figures
                echo=TRUE,                        # shows r-code
                message=FALSE,                    # suppresses library outputs
                warnings=FALSE,                   # suppresses library outputs
                tidy=TRUE,  # prevents the source code from running off a pdf page
                dev='pdf')                        # pdf device
```

#Introduction

This document provides the r-code for the white plague disease model described on the article.The model address to the objectives of this study: (1) to evaluate the annual patterns of WP disease, and (2) to identify and quantify the influence of environmental and BIOLOGICAL factors on WP disease prevalence using Bayesian inference. We use rstan for statistical inference (http://mc-stan.org/about/). 

#Data
The white plague prevalence data and corresponding environmental and biological variables used in this model can be downloaded from xxxxxxx. The file "whiteplague_working.RData" needs to be placed in a folder called "Data".
"

#Model

##Load libraries and create directory for results
```{r}
# Date format - forms part of names of created files or graphs
today <- format(Sys.time(), "%Y.%m.%d")
set.seed(2017)

# load libraries
library(formatR)
library(StanHeaders)
library(rstan)
rstan_options(auto_write = TRUE) # automatically save the compilied Stan program
options(mc.cores = parallel::detectCores()) # execute multiple Markov chains in paral

library(MCMCpack) # rwish function
library(modeest)
library(IDPmisc)
library(ggplot2) # for graphs
library(brms) # To include auto-correlation structures in the Stan model/new package Sep 2017 
library(brm)
library(DHARMa) #to extract posterior
```

## Load data
```{r}
rm(list=ls(all=TRUE))
# Load settings
source("r-code/00_settings.r")

load(file="data/Whiteplague_workingdata.Rdata")
head(Whiteplague_workingdata)
dim(Whiteplague_workingdata)
```

#aggregate working_data file per year
```{r}
#Make a new data fream only with annual mean values for each variables, in this way we can join biological variables (only were collected once a year) with environmental variables (here as mean per year).
Whiteplague_workingdata_Annual <-aggregate(Whiteplague_workingdata, by=list(Whiteplague_workingdata$Station_code,Whiteplague_workingdata$Year), 
  FUN=mean, na.rm=TRUE)

Whiteplague_workingdata_Annual


#delete columns without values (NA)
Whiteplague_workingdata_Annual$Station_name <- NULL
Whiteplague_workingdata_Annual$Station_code <- NULL
Whiteplague_workingdata_Annual$Reef.Type <- NULL
Whiteplague_workingdata_Annual$Month <- NULL
Whiteplague_workingdata_Annual$Month_12 <- NULL
Whiteplague_workingdata_Annual$Month_number <- NULL
Whiteplague_workingdata_Annual$Season <- NULL

#Check final data frame
Whiteplague_workingdata_Annual
dim(Whiteplague_workingdata_Annual)
```



## Data preparation
```{r}
df <- Whiteplague_workingdata_Annual[,c("Ntot","ill","Plague","Chl_sca","Oxy_sca","Sal_sca","Turb_sca","Depth_tran_sca","Temp_sca","DHW_sca", "Sand", "Coral", "Sponges","Macroalgae", "Cyanobacteria")]

df$healthy <- df$Ntot-df$ill
df$Year   <- as.factor(Whiteplague_workingdata_Annual$Year)
df$Site    <- as.factor(as.numeric(Whiteplague_workingdata_Annual$Group.1))
df$Obsid   <- as.factor(1:nrow(df))
df$SiteYear <- as.factor(as.numeric(as.factor(paste(df$Year,df$Site,sep="-"))))

df$Site_index <- as.integer(df$Site)
df$Year_index=as.integer(df$Year)
df$Obsid_index=as.integer(df$Obsid)
df$SiteYear_index=as.integer(as.numeric(as.factor(paste(df$Year,df$Site,sep="-"))))

Environment <- df[,c("Chl_sca","Oxy_sca","Sal_sca","Turb_sca","Depth_tran_sca","Temp_sca","DHW_sca","Sand", "Coral", "Sponges","Macroalgae", "Cyanobacteria")]
```


## Model definition
```{r}
dataList = list(
  TotalObservedcolonies = df$Ntot, #Total observed colonies Shallow (2 years) Deep (4 years)
  Diseasedcolonies = df$ill, # Total numer of disease colonies Shallow (2 years) Deep (4 years)
  nWP = nrow(df), #White Plague prevalence total number of rows
  observedWpPrevalence = df$Plague, # Total Prevalence of White Plague Shallow (2 years) Deep (4 years)
  
  Environment = Environment,
  nEnvi = ncol(Environment),
  
  SiteYear = df$SiteYear_index,
  nSiteYear =length(unique(df$SiteYear_index))
)
```


## Model definition
```{r define model}
# Define model
model.Wplague <- '
data {
// counters
int<lower=0> nWP;                      // number of White Plague cases, int=integer  and non negative >0
int<lower=1> nEnvi;                    // number of environmental variables

int<lower=1> nSiteYear;               // number of sites-month combinations
int<lower=1,upper=130> SiteYear[nWP];  // sites-month combinations

matrix[nWP, nEnvi] Environment;

// response
int<lower=0> TotalObservedcolonies [nWP]; // Total number of coral colonies (cases)
int<lower=0> Diseasedcolonies [nWP]; // number of diseased coral colonies (successes)
}
parameters {
real b0; // intercept
// Fixed effects
vector[nEnvi] parEnvironment;   // Environmental-level

// Random effects
real<lower=0> sigmalev_1;
vector[nSiteYear] eta_1;
}



transformed parameters {
vector[nWP] RealDisease;
vector[nSiteYear] siteyear_raw;

siteyear_raw  = eta_1 * sigmalev_1;

RealDisease =  Environment * parEnvironment + siteyear_raw[SiteYear];
} //closes transformed parameter




model{
// Fixed effects priors
b0             ~ normal(0, 10);
parEnvironment ~ normal(0, 10);

//Random effects priors
eta_1 ~ normal(0, 100);

sigmalev_1 ~ uniform(0,100);
//sigmalev_1 ~ student_t(1,0,1);
//sigmalev_1 ~ cauchy(0,2.5);
//sigmalev_1 ~ inv_gamma(0.001, 0.001);

Diseasedcolonies ~ binomial_logit(TotalObservedcolonies,RealDisease);
}



generated quantities{ //predictive inference
real DiseasePredictions[nWP]; //vector to store predictions
for(i in 1:nWP) DiseasePredictions [i] = binomial_rng(TotalObservedcolonies[i],inv_logit(RealDisease[i]));

}
'
```


## Running the model

```{r model_run,eval=FALSE}
Nchains=4
time <- proc.time()
stan.white <- stan(model_code = model.Wplague, data = dataList, iter = 30000, chains = Nchains,refresh=100)
(proc.time() - time)/60/60
```

```{r}
NvarEnvironment <- ncol(Environment)
printVarEnvironment <- paste(rep("parEnvironment[",NvarEnvironment),1:NvarEnvironment,rep("]",NvarEnvironment),sep="")

print(stan.white,pars=c("b0",printVarEnvironment),probs=c(.1,.5,.9))
plot(stan.white,pars=c("b0",printVarEnvironment))


#plotting the posterior distribution for the parameters
post_beta<-As.mcmc.list(stan.white,pars=c("b0",printVarEnvironment))
plot(post_beta)
#traceplot(post_beta, inc_warmup = FALSE)
```




```{r}
posterior<-extract(stan.white,inc_warmup=F,permute=F)

Ntot <- Whiteplague_workingdata$Ntot
observedWhitePlague <- Whiteplague_workingdata$ill

Nchains=Nchains # number of chains, see model definition
modelpredictionsChains <- paste(rep("modelPredictions[,,",Nchains),1:Nchains,rep("]",Nchains),sep="")


# FITTED PREDICTED
modelPredictions <- NULL
modelPredictions<-apply(posterior, MARGIN=1:2, FUN = function(draw) draw[grepl("^RealDisease[\\[]",names(draw))])


# combine 4 chains
predictionsLinearWhitePlague <- NULL
assign("predictionsLinearWhitePlague",eval(parse(text=paste("cbind(",paste(modelpredictionsChains, collapse = ","),")",sep=""))))

predictionsResponseMedianWhitePlague <- inv.logit(as.vector(apply(predictionsLinearWhitePlague, 1, median)))*Ntot

# SIMULATED
modelPredictions <- NULL
modelPredictions<-apply(posterior, MARGIN=1:2, FUN = function(draw) draw[grepl("^DiseasePredictions[\\[]",names(draw))])

# combine 4 chains
simulatedPredictedWhitePlague <- NULL
assign("simulatedPredictedWhitePlague",eval(parse(text=paste("cbind(",paste(modelpredictionsChains, collapse = ","),")",sep=""))))


WhitePlagueDHARMaResponse <- createDHARMa(simulatedResponse = simulatedPredictedWhitePlague,
  observedResponse = observedWhitePlague,fittedPredictedResponse = predictionsResponseMedianWhitePlague,integerResponse = T)


plotSimulatedResiduals(simulationOutput = WhitePlagueDHARMaResponse)
```



#Save the model run results
```{r}
save(stan.white, file =paste("data/2018.02.23c_Whiteplague_model_result.Rdata",sep=""))
```




###Test overdispersion
```{r}
#testOverdispersion(simulationOutput = WhitePlagueDHARMaResponse)
```
###Test zero inflation
```{r}
#testZeroInflation(WhitePlagueDHARMaResponse)
```

#Save the model run results
```{r}
save(stan.white, file =paste("data/2017.11.21a_Whiteplague_model_result.Rdata",sep=""))
```

# Session information for model run
```{r}
devtools::session_info()
```

#Plotting the model
```{r ploting from stan to ggmcmc}

print(stan.white)



plot(stan.white)

 str(stan.white)

library(ggmcmc)
library(coda)

#This is to Install Stan Caterpillar StanCat to build caterpillar graphs from Stan
#devtools::install_github('christophergandrud/StanCat')
StanCat::stan_caterpillar(stan.white, pars = 'v_raw\\[.*\\]')

#To use ggmcmc#the rstan::extract, specify that is extracted from rstan procedure and not from dpylr
library(plyr)
#this one extracts everything but is not plotting histograms properly
s <- rstan::extract(stan.white, inc_warmup=FALSE, permuted=FALSE) 
#this one extracts only some functions

s <- mcmc.list(mcmc(s[,1,-4]), mcmc(s[,2,-4]), mcmc(s[,3,-4]), mcmc(s[,4,-4]))
S <- ggs(s) 
ggmcmc(S)
ggs_traceplots(S)
ggs_caterpillar(S)


#for specific parameters. But is not runing
#s <- rstan::extract(stan.white, pars = 'v_raw', permuted = TRUE)$parEnvironment 
#s <- do.call(mcmc.list, alply(s[,,-4], 2, mcmc))

```

#Check model output with Shiny
```{r extract with shiny}
library(shiny)
library(shinystan)

launch_shinystan(stan.white)
```


#Save r markdown file
```{r-To save the r markdown file}

save(dataList, stan.white, file = paste("Figures/", today, "_fit.fullModel.pdf",
sep = ""))
save(dataList, file = paste("Figures/", today, "_fit.fullModel_data.Rdata",
sep = ""))
library(lattice)
trellis.device("pdf", file = paste("Figures/", today, "_Output_rstan.pdf",
sep = ""), width = 14, height = 7)
plot(stan.white)
dev.off()


library(lattice)
trellis.device("pdf")
file= paste("Figures/",today,"_Output_rstan.pdf",sep="2-2016", width=14, height=7)
plot(stan.white)
```

##Cauchy and normal distribution test

```{r}
#Normal distribution
x <- seq(-10, 10, length=100)
hx <- dnorm(x)

degf <- c(1, 3, 8, 30)
colors <- c("red", "blue", "darkgreen", "gold", "black")
labels <- c("df=1", "df=3", "df=8", "df=30", "normal")

plot(x, hx, type="l", lty=2, xlab="x value",
  ylab="Density", main="Comparison of t Distributions")

for (i in 1:4){
  lines(x, dt(x,degf[i]), lwd=2, col=colors[i])
}

legend("topright", inset=.05, title="Distributions",
  labels, lwd=2, lty=c(1, 1, 1, 1, 2), col=colors)

```

#cauchy

```{r}
#Cauchy distribution

x <- seq(-10, 10, length=100)
hi <- dcauchy(x)

degf <- c(1, 3, 8, 30)
colors <- c("red", "blue", "darkgreen", "gold", "black")
labels <- c("df=1", "df=3", "df=8", "df=30", "cauchy")

plot(x, hx, type="l", lty=2, xlab="x value",
  ylab="Density", main="Comparison of t Distributions")

for (i in 1:4){
  lines(x, dt(x,degf[i]), lwd=2, col=colors[i])
}

legend("topright", inset=.05, title="Distributions",
  labels, lwd=2, lty=c(1, 1, 1, 1, 2), col=colors)
```

